### 2016-12-02 paper/main.tex
added paper skeleton

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{emnlp16}
\usepackage{graphicx}
\usepackage[table,xcdraw]{xcolor}

\newcommand{\ms}[1]{{\color{cyan}\{\textit{#1}\}$_{ms}$}}

\title{Benchmarking Python Deep Learning Frameworks\\ on Language Modeling }
% feel free to make this title much much better!
\emnlpfinalcopy

\author{Lucy Lin, George Mulcaire \& Maarten Sap
\\University of Washington}
\date{December 2016}

\begin{document}

\maketitle

\begin{abstract}
Abstraaaact
\end{abstract}

\section{Intro}

\subsection{Language Modeling}

\subsection{Data \& Preprocessing}

\section{Experiments}

\section{Conclusion}
\bibliography{references}
\bibliographystyle{emnlp16}
\end{document}

### 2016-12-02 tensorflow/main.py

@@ -38,24 +38,40 @@ args = p.parse_args()
 log(args)
 
 def load_data(path):
-  pkl = path.replace(".txt",".pkl")    
+  # path is presumably train.txt
+  pkl = path.replace(".txt",".pkl")
   
   if os.path.isfile(pkl):
     with open(pkl,"rb") as f:
-      docs = pickle.load(f)
-      assert len(docs) == 1191848, pkl
+      train = pickle.load(f)
+      assert len(train) == 1072664, pkl
   else:
     with open(path) as f:
-      docs = [[int(w) for w in l.split(" ")] for l in f.readlines()]
+      train = [[int(w) for w in l.split(" ")] for l in f.readlines()]
+      print(path,len(train))
     with open(pkl,"wb+") as f:
-      pickle.dump(docs,f)
+      pickle.dump(train,f)
+
+  # val path
+  val_path = path.replace("train.txt","val.txt")
+  val_pkl = val_path.replace(".txt",".pkl")
+  if os.path.isfile(val_pkl):
+    with open(val_pkl,"rb") as f:
+      val = pickle.load(f)
+      assert len(val) == 119184, val_pkl
+  else:
+    with open(val_path) as f:
+      val = [[int(w) for w in l.split(" ")] for l in f.readlines()]
+      print(path,len(val))
+    with open(val_pkl,"wb+") as f:
+      pickle.dump(val,f)
   
-  max_seq_len = max(map(len,docs))
-  vocab_size = max(map(max,docs)) + 1 # cause 0 indexed
+  max_seq_len = max(map(len,train+val))
+  vocab_size = max(map(max,train+val)) + 1 # cause 0 indexed
   assert vocab_size == VOCAB_SIZE , "Vocab size mismatch "+\
     str(vocab_size)+" should be "+str(VOCAB_SIZE)
   
-  return docs,max_seq_len,vocab_size
+  return train,val,max_seq_len,vocab_size
 
 def yield_batches(docs,args):
   """Yields `args.batch_size`-sized batches, padding with EOS tokens."""
@@ -83,20 +99,13 @@ def yield_batches(docs,args):
 
 def train(model_path, data_path):
   """Trains the language model, testing on a validation set every """
-  docs,max_seq_len,vocab_size = load_data(data_path)
+  train,val,max_seq_len,vocab_size = load_data(data_path)
   args.max_seq_len = max_seq_len
   args.vocab_size = vocab_size
-
-  # splitting docs into training, validation (10% val)
-  lim = len(docs)//10
[…]

### 2016-12-02 paper/main.tex
wrote some of the paper a little bit

@@
 
-\section{Intro}
+\section{Introduction}
+In recent years, deep learning approaches have exploded in popularity due to their performance gains over traditional systems on many tasks.
+Several deep learning frameworks have been released in response to this, each claiming niche improvements over others, but it is not necessarily clear which frameworks are better suited for which types of applications.
+Previous comparisons may now be out-of-date (older versions; lack of newer frameworks) or were run against simpler tasks (which might have hidden other interesting performance characteristics).
 
+In this project, we benchmark deep learning frameworks on a complex natural language processing task and rank them based on different performance metrics.
+Specifically, we look at the following frameworks:
+\begin{itemize}
+	\item TensorFlow \cite{tensorflow}
+	\item Theano \cite{theano}
+	\item DyNet \cite{dynet}
+\end{itemize}
+In natural language processing (NLP), recurrent neural networks (RNNs) are of particular interest because they make use of the sequential nature of language.
+Benchmarks on RNNs have been run before, but most tasks were benchmarked on toy tasks which do not necessarily reflect performance on real data.
+For instance, GitHub user glample \cite{glample} generated random data to run through their RNN, which is not realistic.
+In NLP, we usually turn words into vocabulary-sized one-hot vectors and then embed them (i.e.
+dimensionality reduction) as hidden-sized vectors.
+Sparsely learning the embedding matrix, which is usually in the order of $[20,000 \times 200]$, is likely to severely impact performance.
+
+\section{Background}
 \subsection{Language Modeling}
+The task of language modeling is about estimating probabilities of particular sequences of words.
+It is used as an invaluable component in various real-world applications such as speech recognition, statistical machine translation, etc.
+More formally, if $s=\{START,x_1,x_2,...,x_n,STOP\}$ is an n-word sentence, language modelling will try to estimate, for each word $x_i$
+its probability given its history: $p(x_i|START,x_1,x_2,...,x_{i-1})$. 
+Historically, this task can be done rather simply using an l-th order Markov model, where $p(x_i|START,x_1,x_2,...,x_{i-1}) = p(x_i|x_{i-l},x_{i-l+1},...,x_{i-1})$.
+However, the Markov assumption does not need to be made with recurrent neural networks, as they have the capabilities to encode much longer histories.
+With RNNs, we normalize (using softmax) the output of a nonlinear function $f\left(p(x_i|START,x_1,x_2,...,x_{i-1})\right) =softmax\left(f(x_i,START,x_1,x_2,...,x_{i-1})\right)$.
 
-\subsection{Data \& Preprocessing}
+\subsection{Frameworks}
+\ms{TODO: talk about each framework a little bit more?}
 
 \section{Experiments}
+\subsection{Data \& Preprocessing}
+We chose to focus on the Los Angeles Times subset from 2009 of the GigaWord corpus \cite{gigaword}.
+Each of these documents is a news article, so we used NLTK’s \cite{nltk} sentence splitter on the articles.
+Every sentence is then tokenized using NLTK’s \verb!casual_tokenize!
+function, limiting ourselves to sentences that are $<100$ tokens long.
+In order to limit overfitting, we replace all words that occurred less than $150$ times by a special \verb!OOV!
+symbol (this allows the model to be more robust when encountering previously unseen words).
+We ended up with $1,191,848$ sentences, $27,269,856$ total word tokens and a vocabulary size of $35,642$.
 
+\section{Benchmarks}
+\section{Discussion}
 \section{Conclusion}

### 2016-12-06 poster/poster.tex
added poster file and some graphics

\documentclass[paperwidth=118cm,paperheight=84cm,portrait,margin=8em,fontscale=0.3]{baposter}
%\usepackage{acl-hlt2011}
%84 cm wide x 118 cm high
\usepackage{times}
\usepackage{svg}
\usepackage{color}
\usepackage{multirow}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{setspace}

%%30in X 40in
%available sizes: (24 X 38?) (36X48)
%\usepackage{array}
\usepackage{relsize}		% For \smaller
\usepackage{url}			% For \url
\usepackage{epstopdf}	% Included EPS files automatically converted to PDF to include with pdflatex

\newcommand{\compresslist}{
	\setlength{\itemsep}{1pt}
	\setlength{\parskip}{0pt}
	\setlength{\parsep}{0pt}
}
\newenvironment{itquote}
{\begin{quote}\begin{itshape}\begin{small}	\begin{center}\setlength{\parskip}{0.0em}\setlength{\parsep}{0.0em}}
{\end{center}\end{small}\end{itshape}\end{quote}}

\graphicspath{{./}}	% Root directory of the pictures 
\tracingstats=2			% Enabled LaTeX logging with conditionals
%%% Color Definitions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\definecolor{uwPurple}{HTML}{4b2e83}
%\definecolor{uwPurple}{HTML}{3b3e72}
\definecolor{uwGold}{RGB}{232,211,162}
\definecolor{uwMetallicGold}{RGB}{145,123,76}

\begin{document}
\typeout{Poster rendering started}

%%% Setting Background Image %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\background{
}

\begin{poster}{
	%grid=true,
	columns=3,
	colspacing=2em,
	headerheight = 0.17\textheight ,
	%eyecatcher=false, 
	borderColor=uwMetallicGold,
	headerColorOne=uwPurple,
	headerColorTwo=uwPurple,
	headerFontColor=white,
	% Only simple background color used, no shading, so boxColorTwo isn't necessary
	boxColorOne=white,
	boxColorTwo=uwPurple,
	headerborder=closed,
	headershape=rounded,
	headerfont=\Large\sf\bf,
	textborder=none,
	bgColorOne=uwGold,
	bgColorTwo=uwPurple,
	background=plain,
	headerborder=open,
	boxshade=plain
}
{
%Eye Catcher, empty if option eyecatcher=no
{
	\begin{tabular}{l}
			%\includegraphics[scale=.6]{graphics/pnt5in.png}
			\includegraphics[width=2.5in]{graphics/UWlogo.pdf} \\
			%\includegraphics[scale=.6]{graphics/pnt5in.png}
			%\includegraphics[scale=.6]{graphics/pnt5in.png}
	\end{tabular}
	%\includegraphics[scale=1.7]{graphics/pnt5in.png}
}
}
{ % GUIDE: GUI for Document Exploration
\textbf{$\spadesuit$ SPADE -- SPan Annotated Document Exploration $\spadesuit$}
}
{
\vspace*{0.8em}
\textbf{Lucy Lin, George Mulcaire \& Maarten Sap}\\

%\vspace*{0.5em}
CSE 512 Data Visualization Final Project\\
Computer Science and Engineering, University of Washington

 \vspace*{6pt}
}
{
	\begin{tabular}{r}
			\includegraphics[width=1.80in]{graphics/CSElogo.png}\includegraphics[scale=.7]{graphics/pnt5in.png}
	\end{tabular}
}

%For poster contributors, the format is as follows.  Posterboards are 3 feet (90cm) high, 4 feet (120cm) wide self-standing boards that will be placed on top of conference tables. Materials for attaching posters to the boards (double-sided tape, pushpins, clips) will be provided.  Poster authors should finish setting up their posters by the end of the morning break, at the latest.  Also, please note the short "Poster Teasers" in the schedule before the poster session, where you will have 60 seconds (strictly timed!) to give a quick overview that will get people excited about coming to see your poster.
%
%For everyone, please bear in mind that a central goal of the workshop is to create discussion between computational linguists and clinical psychologists.  Therefore we would ask that you please formulate your presentations or posters for a diverse audience, rather than the more typical, purely technical ACL crowd.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%Column 0

%% Box 0
\headerbox{Deep Learning and NLP}{name=intro1,column=0,row=0.00, span=1}{
\color{black}
\vspace*{5pt}
\textbf{NLP for NLP}
\begin{itemize}\addtolength{\itemsep}{-0.25\baselineskip}
\item nlp is natural
\item the natural is about the processing, not the language
\end{itemize}\vspace*{5pt}

\begin{tabular}{cc}
\includegraphics[width=.24\columnwidth]{graphics/python.png}
\hspace*{20pt} & \hspace*{-15pt}\parbox{0.62\textwidth}{
\vspace*{-40pt}\textbf{Deep learning}
\begin{itemize}\addtolength{\itemsep}{-0.5\baselineskip}
\item wow
\item neurons
\item just a function
\item such deep
\vspace*{-5pt}
\end{itemize}}
\end{tabular}
}

%% Box 2
\headerbox{Dataset}{name=intro1,column=0,row=0.40, span=1}{
\color{black}
English GigaWord corpus (LDC), LA Times from 2009 ? \vspace*{10pt}
\\
\begin{tabular}{cc}
\parbox{0.45\textwidth}{
\vspace*{-0pt}
\includegraphics[width=.24\columnwidth]{graphics/newspaper.pdf}
\includegraphics[width=.24\columnwidth]{graphics/LAtimes.jpeg}

}
&
\hspace*{-10pt}
\parbox{0.50\textwidth}{
\textbf{Data}
\begin{itemize}\addtolength{\itemsep}{-0.5\baselineskip}
\item Split into sentences then words using NLTK
\item $1,191,848$ sentences
\item $27,269,856$ words (``tokens'') total
\item $35,642$ distinct words in the end
%\vspace*{-5pt}
\end{itemize}}
\end{tabular}
}

%% Box 2
\headerbox{Hyper params?}{name=intro1,column=0,row=0.75, span=1}{
\color{black}
??
\vspace{.4pt}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% COLUMN 1

%% BOX 0
\headerbox{Logos \& Graphs}{name=results,column=1,row=0.00, span=2}{
\includegraphics[width=.24\columnwidth]{graphics/TensorFlow.png}
\includegraphics[width=.24\columnwidth]{graphics/Theano.jpeg}
\includegraphics[width=.24\columnwidth]{graphics/DyNet.pdf}
}

%% BOX 1
%\headerbox{Differential Word Clouds}{name=results,column=1,row=0.45, span=1}{
%\begin{center}
%\includegraphics[scale=.5]{graphics/wordcloud1.pdf}
%\includegraphics[scale=.5]{graphics/wordcloud2.pdf}
%\end{center}
%
%Something about the method for creating them ?
%}

%% Box 3
\headerbox{What else?}{name=methods,column=1,row=0.69, span=1}{
\color{black}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% COLUMN 2

%% BOX 0
%\headerbox{Search Results}{name=results,column=2,row=0, span=1}{
%\begin{center}
%\includegraphics[scale=.5]{graphics/searchResults.JPG}
%
%\end{center}
%\begin{itemize}\addtolength{\itemsep}{-0.5\baselineskip}%\addtolength{\itemindent}{-5pt}
%\item Search for words or phrases in the document text
%\item Filter on frames to examine relationships between the language used and the frame evoked
%\item Search via the wordcloud, the search bar, or by clicking on frames in search results
%\end{itemize}
%}

%% BOX 1
%\headerbox{Document Viewer}{name=results,column=2,row=0.47, span=1}{
%\begin{center}
%\includegraphics[scale=.4]{graphics/documentView.png}
%\end{center}
%
%The document viewer enables a user to examine a specific text with
%clearly marked frame annotations.
%}

%% BOX 2
\headerbox{How many GPUs should I buy?}{name=results,column=2,row=0.69, span=1}{
}

\end{poster}

\end{document}

### 2016-12-08 tensorflow/main.py
various mini tweaks

@@ -107,6 +107,8 @@ def train(model_path, data_path):
   val_b = [b for b in yield_batches(val,args)]
   train_b = [b for b in yield_batches(train,args)]
 
+  log("{} val, {} training batches".format(len(val_b),len(train_b)))
+
   # Building models/graphs
   m = LanguageModel(args)
   
@@ -131,7 +133,7 @@ def train(model_path, data_path):
         *cost_history[-4:]))
       save_model(args.train,m)
       break
-    
+    sys.stdout.flush()
 
 def main(args):
   if args.train and args.data_path:
@@ -150,3 +152,4 @@ if __name__ == "__main__":
   time_d = time.time() - start_time
   log("Done! [That took {}]".format(timedelta(seconds=time_d)))
 
+  sys.exit(0)

### 2016-12-09 poster/poster.tex
added a bunch on the poster

@@
 \usepackage{latexsym}
-\usepackage{amsmath}
+\usepackage{amsmath,amssymb}
 \usepackage{setspace}
@@
 			%\includegraphics[scale=.6]{graphics/pnt5in.png}
-			\includegraphics[width=2.5in]{graphics/UWlogo.pdf} \\
+			\includegraphics[width=2in]{graphics/UWlogo.pdf} \\
 			%\includegraphics[scale=.6]{graphics/pnt5in.png}
@@
 { % GUIDE: GUI for Document Exploration
-\textbf{$\spadesuit$ SPADE -- SPan Annotated Document Exploration $\spadesuit$}
+\textbf{Benchmarking Python Deep Learning \\Frameworks on Language Modeling }
 }
 {
-\vspace*{0.8em}
+\vspace*{7pt}
 \textbf{Lucy Lin, George Mulcaire \& Maarten Sap}\\
@@
 %\vspace*{0.5em}
-CSE 512 Data Visualization Final Project\\
+CSE 550 Systems\\
 Computer Science and Engineering, University of Washington
@@
 	\begin{tabular}{r}
-			\includegraphics[width=1.80in]{graphics/CSElogo.png}\includegraphics[scale=.7]{graphics/pnt5in.png}
+	\includegraphics[scale=.6]{graphics/pnt5in.png}
+	\includegraphics[width=1.4in]{graphics/CSElogo.png}
+	\includegraphics[scale=.7]{graphics/pnt5in.png}
 	\end{tabular}
@@
 \vspace*{5pt}
-\textbf{NLP for NLP}
+\textbf{Deep learning ``revolution'' in natural language processing}
 \begin{itemize}\addtolength{\itemsep}{-0.25\baselineskip}
-\item nlp is natural
-\item the natural is about the processing, not the language
+\item State-of-the-art results are achieved with neural methods \\
+(e.g. machine translation, speech recognition)
+\item Panoply of toolkits started emerging
 \end{itemize}\vspace*{5pt}
 
-\begin{tabular}{cc}
-\includegraphics[width=.24\columnwidth]{graphics/python.png}
-\hspace*{20pt} & \hspace*{-15pt}\parbox{0.62\textwidth}{
-\vspace*{-40pt}\textbf{Deep learning}
+\begin{tabular}{ccc}\hspace*{10pt}
+\includegraphics[scale=.2]{graphics/pnt5in.png}\\
+\includegraphics[width=.24\columnwidth]{graphics/python.jpg}
+ & \parbox{0.35\textwidth}{
+\vspace*{-40pt}\textbf{Python Frameworks}
 \begin{itemize}\addtolength{\itemsep}{-0.5\baselineskip}
-\item wow
-\item neurons
-\item just a function
-\item such deep
-\vspace*{-5pt}
+\item TensorFlow
+\item Theano
+\item DyNet
+\vspace*{0pt}
 \end{itemize}}
+& \parbox{0.2\textwidth}{\vspace*{-40pt}
+\includegraphics[width=40pt]{graphics/TensorFlow.png}\vspace*{3pt}\\
+\includegraphics[width=40pt]{graphics/theano.jpeg}\vspace*{-1pt}\\
+\includegraphics[width=40pt]{graphics/DyNet.pdf}
+}
 \end{tabular}
+\begin{center}
+\textbf{\textit{Which of these is better when tested on a real NLP task:\\ language modeling?}}
+\end{center}
 }
@@
 \headerbox{Dataset}{name=intro1,column=0,row=0.40, span=1}{
-\color{black}
-English GigaWord corpus (LDC), LA Times from 2009 ? \vspace*{10pt}
+\color{black}\vspace*{3pt}
+English GigaWord corpus (LDC), LA Times from 2009 \vspace*{10pt}
 \\
+\hspace*{5pt}
 \begin{tabular}{cc}
-\parbox{0.45\textwidth}{
+\parbox{0.27\textwidth}{
 \vspace*{-0pt}
-\includegraphics[width=.24\columnwidth]{graphics/newspaper.pdf}
-\includegraphics[width=.24\columnwidth]{graphics/LAtimes.jpeg}
+%\includegraphics[width=.24\columnwidth]{graphics/newspaper.pdf}
+\includegraphics[width=.24\columnwidth]{graphics/latimes.jpg}
 
@@
 \hspace*{-10pt}
-\parbox{0.50\textwidth}{
+\parbox{0.6\textwidth}{
 \textbf{Data}
 \begin{itemize}\addtolength{\itemsep}{-0.5\baselineskip}
-\item Split into sentences then words using NLTK
+\item Split into sentences then words \\ using \texttt{NLTK}
 \item $1,191,848$ sentences
@@
 %% Box 2
-\headerbox{Hyper params?}{name=intro1,column=0,row=0.75, span=1}{
+\headerbox{Model specifics}{name=intro1,column=0,row=0.67, span=1}{
 \color{black}
-??
+\begin{itemize}\addtolength{\itemsep}{-0.5\baselineskip}
+\item Words are embedded in the hidden sized space: $\mathbb{N}^{35642} \rightarrow \mathbb{R}^{256}$.
+\item Embedded words are run through the RNN: $w_i = f(w_0^{i-1},\Theta)$.
+\item Cross-entropy loss is used to learn the parameters.
+\end{itemize}
+\begin{center}
+\begin{tabular}{cc}
+\textbf{hyperparam} & \textbf{value} \\\hline
+RNN type & \texttt{LSTM} \\
+hidden size & $256$ \\
+optimizer & \texttt{AdamOptimizer} \\
+learning rate & $0.003$ \\
+batch size & 25 \\
+
+\end{tabular}
+\end{center}
 \vspace{.4pt}
@@
 %% BOX 0
-\headerbox{Logos \& Graphs}{name=results,column=1,row=0.00, span=2}{
+\headerbox{Logos \& Graphs}{name=results,column=1,row=0.00, span=1}{
 \includegraphics[width=.24\columnwidth]{graphics/TensorFlow.png}
-\includegraphics[width=.24\columnwidth]{graphics/Theano.jpeg}
+\includegraphics[width=.24\columnwidth]{graphics/theano.jpeg}
 \includegraphics[width=.24\columnwidth]{graphics/DyNet.pdf}
@@
 %% BOX 0
-%\headerbox{Search Results}{name=results,column=2,row=0, span=1}{
+\headerbox{Search Results}{name=results,column=2,row=0, span=1}{
 %\begin{center}
@@
 %\end{itemize}
-%}
+}
 

### 2016-12-09 tensorflow/main.py

@@ -104,8 +104,8 @@ def train(model_path, data_path):
   args.vocab_size = vocab_size
   
   # Getting batches
-  val_b = [b for b in yield_batches(val,args)]
-  train_b = [b for b in yield_batches(train,args)]
+  val_b = [b for b in yield_batches(val,args)][:10]
+  train_b = [b for b in yield_batches(train,args)][:10]
 
   log("{} val, {} training batches".format(len(val_b),len(train_b)))
 

### 2016-12-11 paper/main.tex
added a bunch on the poster
wrote some of the paper

@@
 \usepackage{emnlp16}
-\usepackage{graphicx}
+\usepackage{graphicx,hyperref}
 \usepackage[table,xcdraw]{xcolor}
@@
 Benchmarks on RNNs have been run before, but most tasks were benchmarked on toy tasks which do not necessarily reflect performance on real data.
-For instance, GitHub user glample \cite{glample} generated random data to run through their RNN, which is not realistic.
+For instance, GitHub user \verb!glample!
+generated random data to run through their RNN \cite{glample}, which is not realistic.
 In NLP, we usually turn words into vocabulary-sized one-hot vectors and then embed them (i.e.
@@
 \subsection{Frameworks}
-\ms{TODO: talk about each framework a little bit more?}
+All three frameworks use a symbolic computation graph, which maps operations and variables to nodes in a graph.
+To perform computation, one has to ``query'' the graph, providing values for variables needed to compute the desired output.
+Figure~\ref{fig:compGraph}\footnote{Taken from \cite{tensorflow}} shows an example of a simple computation graph.
+All three frameworks efficiently handle queries, by only computing nodes that are necessary for the output requested.\ms{Can we confirm that's true?}
+
+\begin{figure}\begin{center}
+\includegraphics[scale=.4]{graphics/tensorflowGraph.png}
+\caption{\label{fig:compGraph}Simple neural computation graph example. \texttt{ReLU} is a rectified linear unit.} 
+\end{center}\end{figure}
+
+All our frameworks run a \texttt{Python} wrapper, with a \texttt{C++} backend.
+\ms{TODO: more spiel?}
+%TensorFlow's computation graph is pure \texttt{Python}, which makes it slower than other frameworks.
 
@@
 
+\subsection{Implementation \& Hyperparameters}
+Table~\ref{tab:hyperparams} summarizes the hyperparameters chosen for the experiments; there was no cross-validation done since the evaluative performance of the model trained is irrelevant.
+\begin{table}\begin{center}
+\begin{tabular}{cc}
+\textbf{hyperparam} & \textbf{value} \\\hline
+RNN type & \texttt{LSTM} \\
+hidden size & $256$ \\
+optimizer & \texttt{AdamOptimizer} \\
+learning rate & $0.003$ \\
+batch size & 25 \\
+\end{tabular}
+\caption{\label{tab:hyperparams}Hyperparameters used in the experiments}
+\end{center}\end{table}
+
+Traditionally, neural models are trained until their performance on a \textit{development} set stops improving.
+Therefore, a full epoch of training (where all the training data is used) is usually followed by a predictive pass through development data.
 \section{Benchmarks}
+\subsection{Time}
+We timed various things.
+\begin{tabular}{c|ccc}
+Time & Dynet & TensorFlow & Theano \\ \hline
+Convergence & & 35h 25' & \\
+Train epoch & & &  \\
+Test epoch & & & \\
+Train batch & & &  \\
+Test batch & & & \\
+\end{tabular}
+
 \section{Discussion}
 \section{Conclusion}
+\newpage
 \bibliography{references}

### 2016-12-12 poster/poster.tex
slightly update outline/notes on poster

@@
 %% Box 3
-\headerbox{What else?}{name=methods,column=1,row=0.69, span=1}{
+\headerbox{Development experience}{name=methods,column=1,row=0.69, span=1}{
 \color{black}
+Implementing stuff sucked/was great/was average
 }
@@
 %% BOX 0
-\headerbox{Search Results}{name=results,column=2,row=0, span=1}{
-%\begin{center}
-%\includegraphics[scale=.5]{graphics/searchResults.JPG}
-%
-%\end{center}
-%\begin{itemize}\addtolength{\itemsep}{-0.5\baselineskip}%\addtolength{\itemindent}{-5pt}
-%\item Search for words or phrases in the document text
-%\item Filter on frames to examine relationships between the language used and the frame evoked
-%\item Search via the wordcloud, the search bar, or by clicking on frames in search results
-%\end{itemize}
+\headerbox{Results}{name=results,column=2,row=0, span=1}{
+\begin{itemize}
+\item Time per epoch (for each system)
+\item Time per batch
+\item Most used function/ most resources consumed by (for each system)
+\item Metrics for that function (possibly others that are shared?)
+\item Time to create computation graph
+\end{itemize}
+
 }
@@
 %% BOX 2
-\headerbox{How many GPUs should I buy?}{name=results,column=2,row=0.69, span=1}{
+\headerbox{Conclusions and Recommendations}{name=results,column=2,row=0.69, span=1}{
+
 }

### 2016-12-13 paper/main.tex
wrote more of the paper

@@
 \subsection{Time}
-We timed various things.
+The first benchmark to look at is how long training takes. Table \ref{tab:timing} summarizes various breakdowns.
+\begin{table}
 \begin{tabular}{c|ccc}
@@
 Convergence & & 35h 25' & \\
-Train epoch & & &  \\
-Test epoch & & & \\
-Train batch & & &  \\
-Test batch & & & \\
+Train epoch & & 6h57' &  \\
+Test epoch & & 7'45'' & \\
+Train batch & & 5'' &  \\
+Test batch & & $<$1'' & \\
 \end{tabular}
+\caption{\label{tab:timing}Total runtime on various subtasks.}
+\end{table}
 
+\subsection{Cuda Profiler}
+We used the CUDA Profiler Toolkit \cite{nvprof} to characterized our implementations, specifically on a simplified task of training on 1 batch.
+
+We profile which methods take the most amount of time, and look into a variety of metrics.
+\begin{itemize}
+\item \verb!achieved_occupancy! - fraction of GPU warps used v.s. total number of warps
+\item \verb!sm_efficiency! - how much of the runtime is spent performing actual computation
+\item \verb!warp_execution_efficiency! - how efficient is execution within a warp; number of registers per thread is a limiting factor.
+\item \verb!warp_nonpred_execution_efficiency! - how efficient is execution within a warp on non-predicated (i.e. non-branching) instructions.
+\item \verb!dram_read_throughput! - read throughput
+\item \verb!dram_write_throughput! - write throughput
+\end{itemize}
 \section{Discussion}
+When looking to use neural network benchmarks, researchers might want to consider various aspects.
+\subsection{Programmer Experience}
+When adopting new software frameworks, it is important to consider the potential learning curve, as well as higher-level advantages of each framework.
+
+\subsection{TensorFlow}
+Pros: easy to use; good support for RNNs and language models; 
+Cons: symbolic computation graph is hard to debug; seems slower than others; 
 \section{Conclusion}
+\section*{Acknowledgments}
+We want to thank Emily Furst for helping us with GPU profiling, specifically understanding the CUDA Profiler documentation.
 \newpage

### 2016-12-13 paper/main.tex
more paper things

@@
 \begin{abstract}
-Abstraaaact
+Neural networks are omnipresent in natural language processing. 
+We benchmark three popular Python frameworks (DyNet, TensorFlow and Theano) on a realistic NLP task: language modeling.
+We find that ...
 \end{abstract}
@@
 \subsection{Implementation \& Hyperparameters}
+In our experiments, we optimized for a standard language modeling objective: \textit{cross-entropy}.
+
 Table~\ref{tab:hyperparams} summarizes the hyperparameters chosen for the experiments; there was no cross-validation done since the evaluative performance of the model trained is irrelevant.
@@
 When adopting new software frameworks, it is important to consider the potential learning curve, as well as higher-level advantages of each framework.
+\subsubsection{DyNet}
 
-\subsection{TensorFlow}
+\subsubsection{TensorFlow}
 Pros: easy to use; good support for RNNs and language models; 
-Cons: symbolic computation graph is hard to debug; seems slower than others; 
+Cons: symbolic computation graph is hard to debug; seems slower than others.
+\subsubsection{Theano}
 \section{Conclusion}
+
 \section*{Acknowledgments}

### 2016-12-13 poster/poster.tex
poster update (dev experience, hardware, function call table template)

@@
 \usepackage{setspace}
+\usepackage{tabularx}
 
@@
 %% BOX 0
-\headerbox{Logos \& Graphs}{name=results,column=1,row=0.00, span=1}{
+\headerbox{Hardware}{name=methods,column=1,row=0.00,span=1}{
+\vspace{0.5em}
+\begin{itemize}\addtolength{\itemsep}{-0.5\baselineskip}
+\item \textbf{CPU}: Intel Core i7-5930K, 6 core, 12 thread (3.5 GHz)
+\item \textbf{Memory}: G.SKILL Ripjaws 4 Series (8 x 8 GB)
+\item \textbf{GPU}: GeForce GTX TITAN X (2 x 12 GB)
+\end{itemize}
+
+TODO: maybe convert into methodology section -- what experiments we ran
+(until convergence, single epoch, single batch)
+}
+
+\headerbox{Logos \& Graphs}{name=results,column=1,row=0.30, span=1}{
 \includegraphics[width=.24\columnwidth]{graphics/TensorFlow.png}
@@
 \includegraphics[width=.24\columnwidth]{graphics/DyNet.pdf}
+
+% placeholdering for more appropriate graph
+Total runtime on various subtasks:
+
+\begin{tabular}{c|ccc}
+Time & Dynet & TensorFlow & Theano \\ \hline
+Convergence & & 35h 25' & \\
+Train epoch & & 6h57' &  \\
+Test epoch & & 7'45'' & \\
+Train batch & & 5'' &  \\
+Test batch & & $<$1'' & \\
+\end{tabular}
 }
@@
 %% Box 3
-\headerbox{Development experience}{name=methods,column=1,row=0.69, span=1}{
-\color{black}
-Implementing stuff sucked/was great/was average
+\headerbox{Development experience}{name=methods,column=1,row=0.65, span=1}{
+\setlength{\extrarowheight}{3pt}
+\begin{tabularx}{\textwidth}{l|X|X}
+\textbf{Framework}  & \textbf{Pros} & \textbf{Cons} \\
+\hline
+TensorFlow   & Easy to use; good support for RNNs and language models &
+    Symbolic computation graph is hard to debug; seems slower than others \\
+\hline
+Theano  & ``Lower level'' operations $\rightarrow$ highly customizable &
+    No built-in support for RNNs (either need to implement or use a wrapper like Keras) \\
+\hline
+DyNet   & TODO   & TODO \\
+\end{tabularx}
 }
@@
 %% BOX 0
-\headerbox{Results}{name=results,column=2,row=0, span=1}{
-\begin{itemize}
-\item Time per epoch (for each system)
-\item Time per batch
-\item Most used function/ most resources consumed by (for each system)
-\item Metrics for that function (possibly others that are shared?)
-\item Time to create computation graph
-\end{itemize}
+\headerbox{Function calls}{name=results,column=2,row=0, span=1}{
+%\begin{itemize}
+%\item Time per epoch (for each system)
+%\item Time per batch
+%\item Most used function/ most resources consumed by (for each system)
+%\item Metrics for that function (possibly others that are shared?)
+%\item Time to create computation graph
+%\end{itemize}
 
+As expected for neural networks, most of the runtime is spent in
+matrix multiplication (variants of \texttt{magma\_lds128\_sgemm\_kernel}).
+
+\vspace{1em}
+
+\begin{tabularx}{\textwidth}{l|r}
+\textbf{Framework}  & \textbf{Percent time spent in matrix multiply} \\
+\hline
+TensorFlow  &  ... \\
+Theano  & ... \\
+DyNet   & ... \\
+\end{tabularx}
+
+TODO: discuss metrics for each.
 }

### 2016-12-13 poster/poster.tex
Updates to poster

@@
 \headerbox{Logos \& Graphs}{name=results,column=1,row=0.30, span=1}{
-\includegraphics[width=.24\columnwidth]{graphics/TensorFlow.png}
-\includegraphics[width=.24\columnwidth]{graphics/theano.jpeg}
-\includegraphics[width=.24\columnwidth]{graphics/DyNet.pdf}
+%\includegraphics[width=.3\columnwidth]{graphics/TensorFlow.png}
+%\includegraphics[width=.3\columnwidth]{graphics/theano.jpeg}
+%\includegraphics[width=.3\columnwidth]{graphics/DyNet.pdf}
+
+Replacement image here?
 
 % placeholdering for more appropriate graph
-Total runtime on various subtasks:
+\begin{center}
+Total runtime on various subtasks, using one GPU:
+\end{center}
 
+\bgroup
+\def\arraystretch{1.1}
+\setlength\tabcolsep{5mm}
 \begin{tabular}{c|ccc}
 Time & Dynet & TensorFlow & Theano \\ \hline
-Convergence & & 35h 25' & \\
-Train epoch & & 6h57' &  \\
-Test epoch & & 7'45'' & \\
-Train batch & & 5'' &  \\
-Test batch & & $<$1'' & \\
+Convergence & 26h10m & 35h25m & (TBD)\\
+Train epoch & x & 6h57m & 6h35m \\
+Test epoch & x & 7m45s & 24m \\
+Train batch & 1s & 5s & x \\
+Test batch & < 1s & $<$1s & x \\
 \end{tabular}
+\egroup
+
 }
@@
 \headerbox{Development experience}{name=methods,column=1,row=0.65, span=1}{
+
 \setlength{\extrarowheight}{3pt}
-\begin{tabularx}{\textwidth}{l|X|X}
+\begin{tabularx}{\textwidth}{l|X|X|X}
 \textbf{Framework}  & \textbf{Pros} & \textbf{Cons} \\
@@
 \hline
-DyNet   & TODO   & TODO \\
+DyNet   & Allows flexible models that change during computation & Less documentation and support; fragile dependencies  \\
 \end{tabularx}
@@
 %% BOX 0
-\headerbox{Function calls}{name=results,column=2,row=0, span=1}{
+\headerbox{Function-level metrics}{name=results,column=2,row=0, span=1}{
 %\begin{itemize}

### 2016-12-13 poster/poster.tex
poster tweaks and images added

@@
 %% Box 2
-\headerbox{Dataset}{name=intro1,column=0,row=0.40, span=1}{
+\headerbox{Dataset}{name=intro1,column=0,row=0.415, span=1}{
 \color{black}\vspace*{3pt}
@@
 %% Box 2
-\headerbox{Model specifics}{name=intro1,column=0,row=0.67, span=1}{
+\headerbox{Model specifics}{name=intro1,column=0,row=0.7, span=1}{
 \color{black}
@@
 %% BOX 0
-\headerbox{Hardware}{name=methods,column=1,row=0.00,span=1}{
+\headerbox{Hardware}{name=hardware,column=1,row=0.00,span=1}{
 \vspace{0.5em}
+\hspace{-15pt}
+\begin{tabular}{cc}
+\parbox{0.85\textwidth}{
 \begin{itemize}\addtolength{\itemsep}{-0.5\baselineskip}
@@
 \item \textbf{GPU}: GeForce GTX TITAN X (2 x 12 GB)
+\end{itemize}\vspace*{35pt}
+} & \hspace*{-10pt}
+\includegraphics[scale=.1]{graphics/titanx.png}\vspace*{-35pt}
+\end{tabular}
+% TODO: maybe convert into methodology section -- what experiments we ran (until convergence, single epoch, single batch)
+}
+\headerbox{Benchmarking strategy}{name=methods,column=1,row=0.1965, span=1}{
+\vspace{5pt}
+\begin{itemize}
+\item \textbf{Timing}:\\convergence, single epoch, single batch
+\item \textbf{GPU Profiling}:\\overall, matrix multiply, specific GPU metrics
 \end{itemize}
-
-TODO: maybe convert into methodology section -- what experiments we ran
-(until convergence, single epoch, single batch)
+\vspace{5pt}
 }
 
-\headerbox{Logos \& Graphs}{name=results,column=1,row=0.30, span=1}{
+\headerbox{Runtime}{name=results,column=1,row=0.415, span=1}{
 %\includegraphics[width=.3\columnwidth]{graphics/TensorFlow.png}
@@
 
-Replacement image here?
-
 % placeholdering for more appropriate graph
+\includegraphics[scale=.07]{graphics/hourglass.jpg}\vspace*{-35pt}
 \begin{center}
-Total runtime on various subtasks, using one GPU:
+\textbf{Total runtime on various subtasks, using one GPU:}
 \end{center}
@@
 \egroup
-
+\vspace{5pt}
 }
@@
 %% Box 3
-\headerbox{Development experience}{name=methods,column=1,row=0.65, span=1}{
-
+\headerbox{Development experience}{name=methods,column=1,row=0.70, span=1}{
 \setlength{\extrarowheight}{3pt}
-\begin{tabularx}{\textwidth}{l|X|X|X}
+\begin{tabularx}{\textwidth}{l|X|XX}
 \textbf{Framework}  & \textbf{Pros} & \textbf{Cons} \\
@@
 Theano  & ``Lower level'' operations $\rightarrow$ highly customizable &
-    No built-in support for RNNs (either need to implement or use a wrapper like Keras)  \\
+    No built-in support for RNNs (unless using wrapper like Keras)  \\
 \hline
@@
 \end{tabularx}
+
 }

### 2016-12-13 poster/poster.tex
poster tweaks and images added (2)

@@
 \begin{itemize}
-\item \textbf{Timing}:\\convergence, single epoch, single batch
-\item \textbf{GPU Profiling}:\\overall, matrix multiply, specific GPU metrics
+\item \textbf{Timing}: \textit{how long does it take to train/test?}\\ till convergence, single epoch, single batch
+\item \textbf{GPU Profiling}: \textit{using the CUDA Profiler}\\overall, matrix multiply, specific GPU metrics
 \end{itemize}

### 2016-12-13 paper/main.tex
Add metrics tables

@@
 All our frameworks run a \texttt{Python} wrapper, with a \texttt{C++} backend.
+Both TensorFlow and DyNet make extensive use of the Eigen library for C++, which provides fast low-level linear algebra for matrix and tensor operations.
+Theano's backend uses the BLAS linear algebra libraries instead.
 \ms{TODO: more spiel?}
@@
 In our experiments, we optimized for a standard language modeling objective: \textit{cross-entropy}.
+The built-in cross-entropy loss functions in each framework were slightly different, which could potentially affect the number of iterations required for convergence.
 
@@
 \caption{\label{tab:timing}Total runtime on various subtasks.}
+\end{table}
+\subsection{Function-level metrics}
+Using NVidia's built-in GPU profiler, \verb!nvprof!, we measured various ways in which the different frameworks utilized the GPU hardware, specifically focusing on the matrix multiply operation implemented in the function \verb!magma_lds128_sgemm_kernel()!.
+Table \ref{tab:metrics} summarizes.
+
+\begin{table}
+\begin{tabular}{c|ccc}
+Metric 										& Dynet & TensorFlow & Theano \\ \hline
+\verb!achieved_occupancy!			&			&					&				\\
+\verb!sm_efficiency!					&			&					&				\\
+\verb!warp_efficiency!					&			&					&				\\
+\verb!warp_nonpred_efficiency!	&			&					&				\\
+\verb!global_hit_rate!					&			&					&				\\
+\verb!local_hit_rate!					&			&					&				\\
+\verb!dram_read_throughput!		&			&					&				\\
+\verb!dram_write_throughput!		&			&					&				\\
+\end{tabular}
+\caption{\label{tab:metrics}Total runtime on various subtasks.}
 \end{table}

### 2016-12-13 poster/poster.tex
Add metrics tables

@@
 \begin{tabularx}{\textwidth}{l|r}
-\textbf{Framework}  & \textbf{Percent time spent in matrix multiply} \\
-\hline
+\textbf{Framework}  & \textbf{Percent time spent in matrix multiply} \\  \hline
 TensorFlow  &  ... \\
@@
 
-TODO: discuss metrics for each.
+\vspace{5mm}
+
+Focusing more closely on this matrix multiply function, we can see how different frameworks make use of the GPU with different levels of efficiency.
+
+\begin{tabularx}{\textwidth}{c|ccc}
+Metric 										& Dynet & TensorFlow & Theano \\ \hline
+\texttt{\detokenize{achieved_occupancy}}			&		.	&		.			&		.		\\
+\texttt{\detokenize{sm_efficiency}}					&		.	&		.			&		.		\\
+\texttt{\detokenize{warp_efficiency}}					&		.	&		.			&		.		\\
+\texttt{\detokenize{warp_nonpred_efficiency}}	&		.	&		.			&		.		\\
+\texttt{\detokenize{global_hit_rate}}					&		.	&		.			&		.		\\
+\texttt{\detokenize{local_hit_rate}}					&		.	&		.			&		.		\\
+\texttt{\detokenize{dram_read_throughput}}		&		.	&		.			&		.		\\
+\texttt{\detokenize{dram_write_throughput}}		&		.	&		.			&		.		
+\end{tabularx}
+
 }

### 2016-12-13 poster/poster.tex
Adding numbers to poster (still need to do dynet)

@@
 As expected for neural networks, most of the runtime is spent in
-matrix multiplication (variants of \texttt{magma\_lds128\_sgemm\_kernel}).
+matrix multiplication. Focusing on the most frequent function call
+(a variant of \texttt{magma\_lds128\_sgemm\_kernel}):
 
@@
 \begin{tabularx}{\textwidth}{l|r}
-\textbf{Framework}  & \textbf{Percent time spent in matrix multiply} \\  \hline
-TensorFlow  &  ... \\
-Theano  & ... \\
+\textbf{Framework}  & \textbf{Percent time spent in top \texttt{magma\_lds128...}} \\  \hline
+TensorFlow  &  68.87\% \\
+Theano  & 49.90\% \\
 DyNet   & ... \\
@@
 
-\begin{tabularx}{\textwidth}{c|ccc}
+\begin{tabularx}{\textwidth}{c|lll}
 Metric 										& Dynet & TensorFlow & Theano \\ \hline
-\texttt{\detokenize{achieved_occupancy}}			&		.	&		.			&		.		\\
-\texttt{\detokenize{sm_efficiency}}					&		.	&		.			&		.		\\
-\texttt{\detokenize{warp_efficiency}}					&		.	&		.			&		.		\\
-\texttt{\detokenize{warp_nonpred_efficiency}}	&		.	&		.			&		.		\\
-\texttt{\detokenize{global_hit_rate}}					&		.	&		.			&		.		\\
-\texttt{\detokenize{local_hit_rate}}					&		.	&		.			&		.		\\
-\texttt{\detokenize{dram_read_throughput}}		&		.	&		.			&		.		\\
-\texttt{\detokenize{dram_write_throughput}}		&		.	&		.			&		.		
+\texttt{\detokenize{achieved_occupancy}}			&		.	&		0.062			&		0.062		\\
+\texttt{\detokenize{sm_efficiency}}					&		.	&		22.58\%			&		16.04\%		\\
+\texttt{\detokenize{warp_efficiency}}					&		.	&		100.0\%			&		99.99\%		\\
+\texttt{\detokenize{warp_nonpred_efficiency}}	&		.	&		99.95\%			&		99.91\%		\\
+\texttt{\detokenize{global_hit_rate}}					&		.	&		0.00\%			&		0.00\%		\\
+\texttt{\detokenize{local_hit_rate}}					&		.	&		0.00\%			&		0.00\%		\\
+\texttt{\detokenize{dram_read_throughput}}*		&		.	&		11.07			&		9.97		\\
+\texttt{\detokenize{dram_write_throughput}}**		&		.	&		411.99			&		503.54		\\
 \end{tabularx}
 
+\vspace{0.5em}
+* GB/s \quad ** MB/s
 }

### 2016-12-13 poster/poster.tex
updating theano in center table

@@
 Test epoch & x & 7m45s & 24m \\
-Train batch & 1s & 5s & x \\
-Test batch & < 1s & $<$1s & x \\
+Train batch & 1s & 5s & 1s \\
+Test batch & < 1s & $<$1s & < 1s \\
 \end{tabular}

### 2016-12-13 paper/main.tex
add dynet numbers

@@
 \begin{itemize}
-\item \verb!achieved_occupancy! - fraction of GPU warps used v.s. total number of warps
+\item \verb!achieved_occupancy! - number of GPU warps used vs. total number of warps available
 \item \verb!sm_efficiency! - how much of the runtime is spent performing actual computation
-\item \verb!warp_execution_efficiency! - how efficient is execution within a warp; number of registers per thread is a limiting factor.
+\item \verb!warp_execution_efficiency! - how efficient is execution within a warp; number of registers per thread is likely a limiting factor for all implementations.
 \item \verb!warp_nonpred_execution_efficiency! - how efficient is execution within a warp on non-predicated (i.e. non-branching) instructions.
-\item \verb!dram_read_throughput! - read throughput
-\item \verb!dram_write_throughput! - write throughput
+\item \verb!dram_read_throughput! - read throughput from RAM to GPU
+\item \verb!dram_write_throughput! - write throughput from GPU to RAM
 \end{itemize}

### 2016-12-13 poster/poster.tex
trying to write some conclusion-y things
add dynet numbers
clean up time breakdown

@@
 As expected for neural networks, most of the runtime is spent in
-matrix multiplication. Focusing on the most frequent function call
-(a variant of \texttt{magma\_lds128\_sgemm\_kernel}):
+matrix multiplication. In every framework, one of the most used function calls is
+a variant of \texttt{magma\_lds128\_sgemm\_kernel}, used for matrix multiplication.
 
@@
 
-\begin{tabularx}{\textwidth}{l|r}
-\textbf{Framework}  & \textbf{Percent time spent in top \texttt{magma\_lds128...}} \\  \hline
-TensorFlow  &  68.87\% \\
-Theano  & 49.90\% \\
-DyNet   & ... \\
+\textbf{Time spent in various functions}
+
+\vspace{1em}
+
+\newcommand{\specialcell}[2][c]{%
+  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
+
+\newcolumntype{Y}{>{\centering\arraybackslash}X}
+
+\begin{tabularx}{\textwidth}{c|lll}
+\centering\textbf{Framework} & Dynet &  TensorFlow & Theano \\  \hline
+\specialcell{matrix multiply \\(\texttt{magma\_lds128...})} & 29.64\%  &  68.87\% & 49.90\% \\
+CUDA memset & 40.13\% & 0.00\% & 0.24\% \\
 \end{tabularx}
@@
 Metric 										& Dynet & TensorFlow & Theano \\ \hline
-\texttt{\detokenize{achieved_occupancy}}			&		.	&		0.062			&		0.062		\\
-\texttt{\detokenize{sm_efficiency}}					&		.	&		22.58\%			&		16.04\%		\\
-\texttt{\detokenize{warp_efficiency}}					&		.	&		100.0\%			&		99.99\%		\\
-\texttt{\detokenize{warp_nonpred_efficiency}}	&		.	&		99.95\%			&		99.91\%		\\
-\texttt{\detokenize{global_hit_rate}}					&		.	&		0.00\%			&		0.00\%		\\
-\texttt{\detokenize{local_hit_rate}}					&		.	&		0.00\%			&		0.00\%		\\
-\texttt{\detokenize{dram_read_throughput}}*		&		.	&		11.07			&		9.97		\\
-\texttt{\detokenize{dram_write_throughput}}**		&		.	&		411.99			&		503.54		\\
+\texttt{\detokenize{achieved_occupancy}}			&		0.062	&		0.062	&		0.062		\\
+\texttt{\detokenize{sm_efficiency}}					&		16.64\%	&		22.58\%	&		16.04\%		\\
+\texttt{\detokenize{warp_efficiency}}					&		100.0\%	&		100.0\%	&		99.99\%		\\
+\texttt{\detokenize{warp_nonpred_efficiency}}	&		100.0\%	&		99.95\%	&		99.91\%		\\
+\texttt{\detokenize{global_hit_rate}}					&		2.88\%	&		0.00\%	&		0.00\%		\\
+\texttt{\detokenize{local_hit_rate}}					&		0.00\%	&		0.00\%	&		0.00\%		\\
+\texttt{\detokenize{dram_read_throughput}}*		&		10.07	&		11.07	 	&		9.97 	\\
+\texttt{\detokenize{dram_write_throughput}}**	&		357.63&		411.99 &		503.54	\\
 \end{tabularx}
@@
 
+Many tradeoffs to consider when selecting a framework: none of them
+win across the board.
+\begin{itemize}\addtolength{\itemsep}{-0.5\baselineskip}
+    \item Performance: see above!
+    \item Simple or complicated installation?
+    \item Is it well-documented?
+    \item Is there a steep learning curve?
+    \item Are there convenient built-ins?
+    \item How difficult is it to debug computation graphs?
+\end{itemize}
 }

### 2016-12-13 poster/poster.tex
added useless figure

@@
 \usepackage{tabularx}
+\usepackage{wrapfig}
 
@@
 Convergence		& 26h10m 		& 35h25m 			& (TBD)\\
-Train epoch & x & 6h57m & 6h35m \\
-Test epoch & x & 7m45s & 24m \\
+Train epoch 		& 2h20m		& 6h57m 			& 6h35m \\
+Test epoch 		& 3m4s			& 7m45s 			& 24m \\
 Train batch 		& 1s 				& 5s 					& 1s \\
-Test batch & < 1s & $<$1s & < 1s \\
+Test batch 			& $<$1s 		& $<$1s 			& $<$1s \\
 \end{tabular}
@@
 
-\vspace{1em}
-
+\begin{center}
 \textbf{Time spent in various functions}
+\end{center}
 
@@
 \begin{tabularx}{\textwidth}{c|lll}
-\centering\textbf{Framework} & Dynet &  TensorFlow & Theano \\  \hline
+\centering\textbf{Function} & Dynet &  TensorFlow & Theano \\  
+\hline
 \specialcell{matrix multiply \\(\texttt{magma\_lds128...})} & 29.64\%  &  68.87\% & 49.90\% \\
+\hline
 CUDA memset & 40.13\% & 0.00\% & 0.24\% \\
@@
 Focusing more closely on this matrix multiply function, we can see how different frameworks make use of the GPU with different levels of efficiency.
+
+\begin{center}
+\textbf{Function-level GPU metrics}
+\end{center}
 
@@
 %% BOX 2
-\headerbox{Conclusions and Recommendations}{name=results,column=2,row=0.69, span=1}{
+\headerbox{Conclusions and Recommendations}{name=results,column=2,row=0.68, span=1}{
 
-Many tradeoffs to consider when selecting a framework: none of them
-win across the board.
+\begin{wrapfigure}{r}{0.4\textwidth}
+\includegraphics[scale=.4]{../paper/graphics/tensorflowGraph.png}
+\end{wrapfigure}
+
+Many tradeoffs to consider when selecting a framework: none of the tested frameworks
+win or lose across the board.
+
 \begin{itemize}\addtolength{\itemsep}{-0.5\baselineskip}
@@
 \end{itemize}
+
+\vspace{0.0mm}
+
 }

### 2016-12-13 paper/main.tex
Updating tables from the poster

@@
 \usepackage{graphicx,hyperref}
+\usepackage{tabularx}
 \usepackage[table,xcdraw]{xcolor}
@@
 Time					& Dynet 		& TensorFlow 	& Theano \\ \hline
-Convergence & & 35h 25' & \\
-Train epoch & & 6h57' &  \\
-Test epoch & & 7'45'' & \\
-Train batch & & 5'' &  \\
-Test batch & & $<$1'' & \\
+Convergence		& 26h10m 		& 35h25m 			& (TBD)\\
+Train epoch 		& 2h20m		& 6h57m 			& 6h35m \\
+Test epoch 		& 3m4s			& 7m45s 			& 24m \\
+Train batch 		& 1s 				& 5s 					& 1s \\
+Test batch 			& $<$1s 		& $<$1s 			& $<$1s \\
 \end{tabular}
@@
 \begin{table}
-\begin{tabular}{c|ccc}
-Metric 										& Dynet & TensorFlow & Theano \\ \hline
-\verb!achieved_occupancy!			&			&					&				\\
-\verb!sm_efficiency!					&			&					&				\\
-\verb!warp_efficiency!					&			&					&				\\
-\verb!warp_nonpred_efficiency!	&			&					&				\\
-\verb!global_hit_rate!					&			&					&				\\
-\verb!local_hit_rate!					&			&					&				\\
-\verb!dram_read_throughput!		&			&					&				\\
-\verb!dram_write_throughput!		&			&					&				\\
+\centering
+\begin{tabular}{c|rrr}
+\textbf{Function} & Dynet &  TensorFlow & Theano \\  
+\hline
+matrix multiply & 29.64\%  &  68.87\% & 49.90\% \\
+\hline
+CUDA memset & 40.13\% & 0.00\% & 0.24\% \\
 \end{tabular}
-\caption{\label{tab:metrics}Total runtime on various subtasks.}
+
+\caption{\label{tab:pcttime}Percentage of runtime spent on the most common matrix multiply call used (a variant of \texttt{\detokenize{magma_lds_128_sgemm_kernel()}}) and on CUDA memset operations.
+These numbers are from training on a single batch.}
 \end{table}
 
+\begin{table*}
+\centering
+\begin{tabular}{c|rrr}
+\textbf{Metric} 										& \textbf{Dynet} & \textbf{TensorFlow} & \textbf{Theano} \\ \hline
+\texttt{\detokenize{achieved_occupancy}}			&		0.062	&		0.062	&		0.062		\\
+\texttt{\detokenize{sm_efficiency}}					&		16.64\%	&		22.58\%	&		16.04\%		\\
+\texttt{\detokenize{warp_efficiency}}					&		100.0\%	&		100.0\%	&		99.99\%		\\
+\texttt{\detokenize{warp_nonpred_efficiency}}	&		100.0\%	&		99.95\%	&		99.91\%		\\
+\texttt{\detokenize{global_hit_rate}}					&		2.88\%	&		0.00\%	&		0.00\%		\\
+\texttt{\detokenize{local_hit_rate}}					&		0.00\%	&		0.00\%	&		0.00\%		\\
+\texttt{\detokenize{dram_read_throughput}} (GB/s)		&		10.07	&		11.07	 	&		9.97 	\\
+\texttt{\detokenize{dram_write_throughput}} (MB/s)	&		357.63&		411.99 &		503.54	\\
+\end{tabular}
+
+\caption{\label{tab:metrics} Various CUDA profiler metrics for the task of training a single batch. Descriptions of these metrics are in section \ref{subsec:cudaprof}.}
+\end{table*}
+
 \subsection{Cuda Profiler}
-We used the CUDA Profiler Toolkit \cite{nvprof} to characterized our implementations, specifically on a simplified task of training on 1 batch.
+\label{subsec:cudaprof}
+We used the CUDA Profiler Toolkit \cite{nvprof} to characterize our implementations, specifically on a simplified task of training on one batch.
 

### 2016-12-13 paper/main.tex
Some discussion changes

@@
 \newcommand{\ms}[1]{{\color{cyan}\{\textit{#1}\}$_{ms}$}}
+\newcommand{\lhl}[1]{{\color{magenta}\{\textit{#1}\}$_{lhl}$}}
 
@@
 	\item TensorFlow \cite{tensorflow}
-	\item Theano \cite{theano}
+	\item Theano \cite{theano}, using the Keras wrapper \cite{keras}
 	\item DyNet \cite{dynet}
@@
 \end{itemize}
+
 \section{Discussion}
 When looking to use neural network benchmarks, researchers might want to consider various aspects.
+
+\subsection{Performance}
+
+\lhl{TODO: a higher-level summary of what we found in the benchmarks section.}
+
 \subsection{Programmer Experience}
-When adopting new software frameworks, it is important to consider the potential learning curve, as well as higher-level advantages of each framework.
-\subsubsection{DyNet}
 
-\subsubsection{TensorFlow}
-Pros: easy to use; good support for RNNs and language models; 
-Cons: symbolic computation graph is hard to debug; seems slower than others.
-\subsubsection{Theano}
+While implementing language models in these frameworks, we found that it is also important to consider the development experience.
+No one framework wins or loses out; instead, there are several tradeoffs to consider.
+We describe some of our experiences below.
+
+\subsubsection{Installation/dependencies}
+
+Unlike a standard Python module which just requires a simple \texttt{pip install}, neural network frameworks are typically dependent on external fast low-level linear algebra libraries to perform efficiently.
+Therefore there is often linking or configuration required to wire things together.
+
+\lhl{TODO: fill in other installation bits}
+
+\subsubsection{Ease of use}
+
+\lhl{TODO: all the things that TensorFlow/DyNet support out of the box.
+comment somewhere about documentation, debuggability}
+
+In contrast, Theano is ``lower-level'' in that it requires that the programmer define the shared variables, propagation steps, cost functions, and so on manually.
+Theano therefore offers a great deal of control and flexibility in model implementation, which is a benefit if one is implementing neural net layers not supported by other frameworks.
+
+However, this also presents a steeper learning curve (despite the extensive documentation) and a much higher prototyping/development cost.
+We originally attempted to implement the Theano language model just using Theano and found the learning/debugging costs to be high.
+Because our language model consists of very standard neural net components, we instead switched to using Keras, which can use Theano under the covers but supplies an easier API with built-in RNN components.
+
 \section{Conclusion}

### 2016-12-14 paper/main.tex
commentssss

@@
 
-\title{Benchmarking Python Deep Learning Frameworks\\ on Language Modeling }
+\title{Benchmarking Python Deep Learning Frameworks\\ for Language Modeling on GPUs}
 % feel free to make this title much much better!
@@
 We benchmark three popular Python frameworks (DyNet, TensorFlow and Theano) on a realistic NLP task: language modeling.
-We find that ...
+\ms{Change if needed:} We find that DyNet is significantly faster on this task. However, performance wasn't the only bottleneck during development, and there are other variables to consider.
 \end{abstract}
@@
 
-All our frameworks run a \texttt{Python} wrapper, with a \texttt{C++} backend.
+All our frameworks run a \texttt{Python} wrapper\ms{TODO: add footnote about which version of Python}, with a \texttt{C++} backend.
 Both TensorFlow and DyNet make extensive use of the Eigen library for C++, which provides fast low-level linear algebra for matrix and tensor operations.
 Theano's backend uses the BLAS linear algebra libraries instead.
-\ms{TODO: more spiel?}
+\ms{TODO: more spiel? E.g. static computation graph (theano, tf) vs dynamic (Dynet)}
 %TensorFlow's computation graph is pure \texttt{Python}, which makes it slower than other frameworks.
@@
 \subsection{Implementation \& Hyperparameters}
+\ms{Add equations/words to explain what a neural LM does.}
 In our experiments, we optimized for a standard language modeling objective: \textit{cross-entropy}.
 The built-in cross-entropy loss functions in each framework were slightly different, which could potentially affect the number of iterations required for convergence.
+\ms{TODO: add github somewhere}
 
@@
 \subsection{Time}
-The first benchmark to look at is how long training takes. Table \ref{tab:timing} summarizes various breakdowns.
+The first benchmark to look at is how long training takes. Table \ref{tab:timing} summarizes various breakdowns. \ms{Summarize results: ie Dynet is the best (so far) (2 sentences perhaps)}.
 \begin{table}
@@
 Table \ref{tab:metrics} summarizes.
+\ms{Summarize results: i.e.
+they're all in the same order of magnitute, doesn't really tell us why DyNet would be much more efficient.
+George TODO: explain the memset thing.}
 
@@
 \section{Discussion}
-When looking to use neural network benchmarks, researchers might want to consider various aspects.
+\ms{When looking to use neural network benchmarks, researchers might want to consider various aspects.}
 
@@
 \lhl{TODO: a higher-level summary of what we found in the benchmarks section.}
+\ms{What do our results mean in term of selecting a framework, be broader in terms of how this is task specific.}
 
@@
 \subsubsection{Ease of use}
-
+\ms{TF pro: Google has money and resources.}
 \lhl{TODO: all the things that TensorFlow/DyNet support out of the box.
@@
 
+\paragraph{Theano} 
 In contrast, Theano is ``lower-level'' in that it requires that the programmer define the shared variables, propagation steps, cost functions, and so on manually.
@@
 \section{Conclusion}
+\ms{Many factors, no clear winner; }
 

### 2016-12-14 paper/main.tex
updating 2.2 frameworks section

@@
 All three frameworks use a symbolic computation graph, which maps operations and variables to nodes in a graph.
-To perform computation, one has to ``query'' the graph, providing values for variables needed to compute the desired output.
-Figure~\ref{fig:compGraph}\footnote{Taken from \cite{tensorflow}} shows an example of a simple computation graph.
-All three frameworks efficiently handle queries, by only computing nodes that are necessary for the output requested.\ms{Can we confirm that's true?}
+To perform a computation, one has to ``query'' the graph by providing values for variables needed to compute the desired output.
+Figure~\ref{fig:compGraph} shows an example of a simple computation graph.
 
@@
 \includegraphics[scale=.4]{graphics/tensorflowGraph.png}
-\caption{\label{fig:compGraph}Simple neural computation graph example. \texttt{ReLU} is a rectified linear unit.} 
+\caption{\label{fig:compGraph}Simple neural computation graph example. (\texttt{ReLU} is a rectified linear unit.) Figure is taken from \protect\cite{tensorflow}.} 
 \end{center}\end{figure}
 
-All our frameworks run a \texttt{Python} wrapper\ms{TODO: add footnote about which version of Python}, with a \texttt{C++} backend.
-Both TensorFlow and DyNet make extensive use of the Eigen library for C++, which provides fast low-level linear algebra for matrix and tensor operations.
-Theano's backend uses the BLAS linear algebra libraries instead.
-\ms{TODO: more spiel? E.g. static computation graph (theano, tf) vs dynamic (Dynet)}
+The frameworks handle computation graphs differently.
+Theano and Tensorflow perform ``static'' declaration of computation graphs: first, the graph architecture is specified, and then data is run through the compiled graph to train or make predictions.
+In contrast, DyNet does ``dynamic'' declaration of computation graphs; that is, the computation graph is defined on-the-fly as operations are executed.
+
+All of the frameworks we tested have Python APIs\footnote{DyNet was originally a C++ framework which now has a Python wrapper; currently its Python wrapper only supports Python 2.
+TensorFlow and Theano are primarily Python frameworks that support both Python 2 and 3.} that wrap C++ or optimized Python code (e.g., numpy).
+
+For speed, the frameworks all use fast low-level linear algebra libraries for matrix and tensor operations.
+Both TensorFlow and DyNet make extensive use of the Eigen library for C++, while Theano uses BLAS instead.
+
 %TensorFlow's computation graph is pure \texttt{Python}, which makes it slower than other frameworks.

### 2016-12-14 paper/main.tex
DyNet updates

--- 2016-12-14 paper/main.tex
@@
 Theano's backend uses the BLAS linear algebra libraries instead.
-\ms{TODO: more spiel? E.g. static computation graph (theano, tf) vs dynamic (Dynet)}
+\ms{TODO: more spiel? E.g. static computation graph (theano, tf) vs dynamic (DyNet)}
 %TensorFlow's computation graph is pure \texttt{Python}, which makes it slower than other frameworks.
@@
 \subsection{Time}
-The first benchmark to look at is how long training takes. Table \ref{tab:timing} summarizes various breakdowns. \ms{Summarize results: ie Dynet is the best (so far) (2 sentences perhaps)}.
+The first benchmark to look at is how long training takes. Table \ref{tab:timing} summarizes various breakdowns. \ms{Summarize results: ie DyNet is the best (so far) (2 sentences perhaps)}.
 \begin{table}
 \begin{tabular}{c|ccc}
-Time					& Dynet 		& TensorFlow 	& Theano \\ \hline
+Time					& DyNet 		& TensorFlow 	& Theano \\ \hline
 Convergence		& 26h10m 		& 35h25m 			& (TBD)\\
@@
 \ms{Summarize results: i.e.
-they're all in the same order of magnitute, doesn't really tell us why DyNet would be much more efficient.
-George TODO: explain the memset thing.}
+they're all in the same order of magnitude, nothing really tells us why DyNet would be much more efficient.} DyNet also spends a large amount of time in the \texttt{CUDA memset} function, unlike the other two frameworks.
+This may relate to DyNet's memory model: the total amount of memory available to forward and backward passes through the computation graph is specified at runtime, and DyNet may manually manage its memory use in a way TensorFlow and Theano do not.
 
@@
 \begin{tabular}{c|rrr}
-\textbf{Function} & Dynet &  TensorFlow & Theano \\  
+\textbf{Function} & DyNet &  TensorFlow & Theano \\  
 \hline
@@
 \begin{tabular}{c|rrr}
-\textbf{Metric} 										& \textbf{Dynet} & \textbf{TensorFlow} & \textbf{Theano} \\ \hline
+\textbf{Metric} 										& \textbf{DyNet} & \textbf{TensorFlow} & \textbf{Theano} \\ \hline
 \texttt{\detokenize{achieved_occupancy}}			&		0.062	&		0.062	&		0.062		\\
@@
 
+\paragraph{DyNet}
+DyNet includes built-in constructors for various NLP-relevant models, including several varieties recurrent neural nets like Tree LSTMs and encoder-decoders.
+THis makes it easy for the programmer to build many common kinds of models.
+DyNet also allows for dynamic computation graph, allowing nodes (representing parameter tensors) to be combined ``on the fly'' to adapt to, for example, data sequences of different lengths.
+While our neural language model used a static computation graph, DyNet made graph construction easy.
+
+DyNet's primary drawback is the relative lack of documentation and support.
+The development team and userbase are both fairly small, so there is less information available to help users debug.
+There is some documentation available online, but it is incomplete.
+
 \paragraph{Theano} 

### 2016-12-14 paper/main.tex

@@
 \includegraphics[scale=.4]{graphics/tensorflowGraph.png}
-\caption{\label{fig:compGraph}Simple neural computation graph example. \texttt{ReLU} is a rectified linear unit.} 
+\caption{\label{fig:compGraph}Simple neural computation graph example. (\texttt{ReLU} is a rectified linear unit.) Figure is taken from \protect\cite{tensorflow}.} 
 \end{center}\end{figure}
 
-All our frameworks run a \texttt{Python} wrapper\ms{TODO: add footnote about which version of Python}, with a \texttt{C++} backend.
-Both TensorFlow and DyNet make extensive use of the Eigen library for C++, which provides fast low-level linear algebra for matrix and tensor operations.
-Theano's backend uses the BLAS linear algebra libraries instead.
-\ms{TODO: more spiel? E.g. static computation graph (theano, tf) vs dynamic (DyNet)}
+The frameworks handle computation graphs differently.
+Theano and Tensorflow perform ``static'' declaration of computation graphs: first, the graph architecture is specified, and then data is run through the compiled graph to train or make predictions.
+In contrast, DyNet does ``dynamic'' declaration of computation graphs; that is, the computation graph is defined on-the-fly as operations are executed.
+
+All of the frameworks we tested have Python APIs\footnote{DyNet was originally a C++ framework which now has a Python wrapper; currently its Python wrapper only supports Python 2.
+TensorFlow and Theano are primarily Python frameworks that support both Python 2 and 3.} that wrap C++ or optimized Python code (e.g., numpy).
+
+For speed, the frameworks all use fast low-level linear algebra libraries for matrix and tensor operations.
+Both TensorFlow and DyNet make extensive use of the Eigen library for C++, while Theano uses BLAS instead.
+
 %TensorFlow's computation graph is pure \texttt{Python}, which makes it slower than other frameworks.

### 2016-12-14 paper/main.tex
Add DyNet install stuff
updating 5.1 performance

@@
 \section{Benchmarks}
+\label{sec:benchmarks}
 \subsection{Time}
@@
 \section{Discussion}
-\ms{When looking to use neural network benchmarks, researchers might want to consider various aspects.}
+
+Selecting a neural network framework involves several considerations. In this section, we describe how performance and various aspects of the programming experience may impact such a selection.
 
@@
 
-\lhl{TODO: a higher-level summary of what we found in the benchmarks section.}
-\ms{What do our results mean in term of selecting a framework, be broader in terms of how this is task specific.}
+As discussed in section \ref{sec:benchmarks}, we found that DyNet was the fastest at performing our specific language modeling task.
+\lhl{TODO: flesh this out with a sentence or two more -- maybe something like ``DyNet seems to make certain optimizations that are advantageous in this setting''}
+
+However, we note that the comparative performance between frameworks is likely dependent on the task, data set, choice of optimizer, and other parameters.
+Therefore, these results may not extrapolate to other settings, and if tuning training or prediction performance is a serious consideration, we recommend benchmarking on the specific task at hand.
 

### 2016-12-14 paper/main.tex
update dynet footnote in 2.2; update theano install in 5.2.1

@@
 
-All of the frameworks we tested have Python APIs\footnote{DyNet was originally a C++ framework which now has a Python wrapper; currently its Python wrapper only supports Python 2.
+All of the frameworks we tested have Python APIs\footnote{DyNet was originally a C++ framework which now has a Python wrapper; currently its Python wrapper seems to only support Python 2.
 TensorFlow and Theano are primarily Python frameworks that support both Python 2 and 3.} that wrap C++ or optimized Python code (e.g., numpy).
@@
 
+\paragraph{DyNet} We encountered some difficulties in installing DyNet.
+The library failed to build with the recommended development version of Eigen, so we had to revert to an earlier commit.
+We also had to use Python 2 for the Python interface, though the documentation says it supports Python 3.
+
+\paragraph{TensorFlow}
+
+\paragraph{Theano} Our experience with Theano installation was relatively straightforward: pip install of relevant modules and specification of the BLAS install location.
+
 \subsubsection{Ease of use}
@@
 There is some documentation available online, but it is incomplete.
+
+\paragraph{TensorFlow} \textless TensorFlow stuff goes here \textgreater
 

### 2016-12-14 paper/main.tex
Neural LM description and github link

@@
 
-All of the frameworks we tested have Python APIs\footnote{DyNet was originally a C++ framework which now has a Python wrapper; currently its Python wrapper seems to only support Python 2.
+All of the frameworks we tested have Python APIs\footnote{DyNet was originally a C++ framework which now has a Python wrapper; currently its Python wrapper seems to most reliably support Python 2.
 TensorFlow and Theano are primarily Python frameworks that support both Python 2 and 3.} that wrap C++ or optimized Python code (e.g., numpy).
@@
 \ms{Add equations/words to explain what a neural LM does.}
+We implemented the same recurrent neural language model in each framework.
+The model consists of a matrix of input word embeddings, which are passed through a RNN to get predicted word embeddings, which are compared via a loss function to a corresponding set of output embeddings.
+All these elements are available in all frameworks (in Theano's case, the RNN is provided by the Keras wrapper).
+Our implementations are made available on GitHub.\footnote{\url{https://github.com/lucylin/neural-benchmarks}}
 In our experiments, we optimized for a standard language modeling objective: \textit{cross-entropy}.
 The built-in cross-entropy loss functions in each framework were slightly different, which could potentially affect the number of iterations required for convergence.
-\ms{TODO: add github somewhere}
 

### 2016-12-14 paper/main.tex
remove neural LM todo
reorganizing / fleshing out sections 4.2 and 4.3

--- 2016-12-14 paper/main.tex
@@
 \end{table}
-\subsection{Function-level metrics}
-Using NVidia's built-in GPU profiler, \verb!nvprof!, we measured various ways in which the different frameworks utilized the GPU hardware, specifically focusing on the matrix multiply operation implemented in the function \verb!magma_lds128_sgemm_kernel()!.
-Table \ref{tab:metrics} summarizes.
-\ms{Summarize results: i.e.
-they're all in the same order of magnitude, nothing really tells us why DyNet would be much more efficient.} DyNet also spends a large amount of time in the \texttt{CUDA memset} function, unlike the other two frameworks.
+
+\subsection{Kernel function usage}
+Using NVidia's built-in GPU profiler, \verb!nvprof!, we first looked into which kernel function calls were most heavily utilized by profiling training on a single batch.
+As can be seen in Table~\ref{tab:pcttime}, a large portion of the runtime is spent on matrix multiplication (specifically, a variant of \texttt{\detokenize{magma_lds_128_sgemm_kernel()}}) for all frameworks; this is expected given how neural networks work.
+
+Unlike the other two frameworks, DyNet also spends a large amount of time in the \texttt{CUDA memset} function.
 This may relate to DyNet's memory model: the total amount of memory available to forward and backward passes through the computation graph is specified at runtime, and DyNet may manually manage its memory use in a way TensorFlow and Theano do not.
@@
 
-\subsection{Cuda Profiler}
+\subsection{CUDA profiler metrics}
 \label{subsec:cudaprof}
-We used the CUDA Profiler Toolkit \cite{nvprof} to characterize our implementations, specifically on a simplified task of training on one batch.
 
-We profile which methods take the most amount of time, and look into a variety of metrics.
+Given the high usage of the matrix multiply operation \verb!magma_lds128_sgemm_kernel()!, we
+then looked into a variety of CUDA profiler metrics for this operation.
+
+We collected data for the following metrics:
 \begin{itemize}
-\item \verb!achieved_occupancy! - number of GPU warps used vs. total number of warps available
-\item \verb!sm_efficiency! - how much of the runtime is spent performing actual computation
-\item \verb!warp_execution_efficiency! - how efficient is execution within a warp; number of registers per thread is likely a limiting factor for all implementations.
-\item \verb!warp_nonpred_execution_efficiency! - how efficient is execution within a warp on non-predicated (i.e. non-branching) instructions.
-\item \verb!dram_read_throughput! - read throughput from RAM to GPU
-\item \verb!dram_write_throughput! - write throughput from GPU to RAM
+\item \verb!achieved_occupancy! -- number of GPU warps used vs.\ total number of warps available.
+\item \verb!sm_efficiency! -- how much of the runtime is spent performing actual computation.
+\item \verb!warp_execution_efficiency! -- how efficient execution is within a warp; note that the number of registers per thread is likely a limiting factor for all implementations.
+\item \verb!warp_nonpred_execution_efficiency! -- how efficient execution is within a warp on non-predicated (i.e. non-branching) instructions.
+\item \verb!dram_read_throughput! -- read throughput from RAM to GPU.
+\item \verb!dram_write_throughput! -- write throughput from GPU to RAM.
 \end{itemize}
+
+Table \ref{tab:metrics} summarizes the results on a single batch.
+We found that all of the measurements were within similar orders of magnitude; nothing at this level directly points to why DyNet performance was considerably faster than the other frameworks.
 

### 2016-12-14 paper/main.tex

@@
 \subsection{Implementation \& Hyperparameters}
-\ms{Add equations/words to explain what a neural LM does.}
+We implemented the same recurrent neural language model in each framework.
+The model consists of a matrix of input word embeddings, which are passed through a RNN to get predicted word embeddings, which are compared via a loss function to a corresponding set of output embeddings.
+All these elements are available in all frameworks (in Theano's case, the RNN is provided by the Keras wrapper).
+Our implementations are made available on GitHub.\footnote{\url{https://github.com/lucylin/neural-benchmarks}}
 In our experiments, we optimized for a standard language modeling objective: \textit{cross-entropy}.
 The built-in cross-entropy loss functions in each framework were slightly different, which could potentially affect the number of iterations required for convergence.
-\ms{TODO: add github somewhere}
 

### 2016-12-14 paper/main.tex
added some tensorflow stuff

@@
 \lhl{TODO: flesh this out with a sentence or two more -- maybe something like ``DyNet seems to make certain optimizations that are advantageous in this setting''}
+DyNet seems to make certain optimizations that are advantageous in this setting, perhaps specifically in the memory management choices it makes.\ms{is this okay?}
 
@@
 
-\lhl{TODO: fill in other installation bits}
-
 \paragraph{DyNet} We encountered some difficulties in installing DyNet.
@@
 \paragraph{TensorFlow}
+TensorFlow provides very extensive installation guidelines.
+It installs in the command line through pip, along with installing the dependencies.
+However, it rather inconveniently requires two different installations for GPU or CPU versions, which can cause issues if not using a virtual environment of some sort.
 
@@
 \subsubsection{Ease of use}
-\ms{TF pro: Google has money and resources.}
 \lhl{TODO: all the things that TensorFlow/DyNet support out of the box.
-comment somewhere about documentation, debuggability}
+comment somewhere about documentation, debuggability\ms{Done?}}
 
@@
 DyNet includes built-in constructors for various NLP-relevant models, including several varieties recurrent neural nets like Tree LSTMs and encoder-decoders.
-THis makes it easy for the programmer to build many common kinds of models.
+This makes it easy for the programmer to build many common kinds of models.
 DyNet also allows for dynamic computation graph, allowing nodes (representing parameter tensors) to be combined ``on the fly'' to adapt to, for example, data sequences of different lengths.
@@
 
-\paragraph{TensorFlow} \textless TensorFlow stuff goes here \textgreater
+\paragraph{TensorFlow} TensorFlow is relatively easy to use, as it comes with a myriad of tutorials and examples.
+RNN support, as well as many other standard models, is relatively cutting-edge.
+Writing the computation graph requires some amount of know-how, since a lot if TensorFlow specific syntax; debugging the symbolic graph is hard since there are no values.
+TensorFlow requires a decent amount of boilerplate code, specifically for mini-batching; the computation graph being static means that sequences either need to be padded, or mini-batches do, which can make computation somewhat cumbersome for RNNs.
+Overall, TensorFlow's large userbase, polyvalence and huge amounts of support from its creator (Google) makes up for its flaws.
 

### 2016-12-14 paper/main.tex
Cleanup, add final dynet results, draft conclusion

@@
 We benchmark three popular Python frameworks (DyNet, TensorFlow and Theano) on a realistic NLP task: language modeling.
-\ms{Change if needed:} We find that DyNet is significantly faster on this task. However, performance wasn't the only bottleneck during development, and there are other variables to consider.
+We find that DyNet is significantly faster on this task. However, performance wasn't the only bottleneck during development, and there are other variables to consider.
 \end{abstract}
@@
 \subsection{Time}
-The first benchmark to look at is how long training takes. Table \ref{tab:timing} summarizes various breakdowns. \ms{Summarize results: ie DyNet is the best (so far) (2 sentences perhaps)}.
+The first benchmark to look at is how long training takes.
+Table \ref{tab:timing} summarizes various breakdowns.
+Overall, DyNet seems to converge the fastest of the three; its epoch time is dramatically shorter, but it also took more epochs to converge.
+
 \begin{table}
@@
 Time					& DyNet 		& TensorFlow 	& Theano \\ \hline
-Convergence		& 26h10m 		& 35h25m 			& (TBD)\\
+Convergence		& 26h21m 		& 35h25m 			& (TBD)\\
 Train epoch 		& 2h20m		& 6h57m 			& 6h35m \\
@@
 As discussed in section \ref{sec:benchmarks}, we found that DyNet was the fastest at performing our specific language modeling task.
-\lhl{TODO: flesh this out with a sentence or two more -- maybe something like ``DyNet seems to make certain optimizations that are advantageous in this setting''}
-DyNet seems to make certain optimizations that are advantageous in this setting, perhaps specifically in the memory management choices it makes.\ms{is this okay?}
+DyNet seems to make certain optimizations that are advantageous in this setting, perhaps specifically in the memory management choices it makes, or in adapting to inputs of different sizes.
 
@@
 \subsubsection{Ease of use}
-\lhl{TODO: all the things that TensorFlow/DyNet support out of the box.
-comment somewhere about documentation, debuggability\ms{Done?}}
 
 \paragraph{DyNet}
-DyNet includes built-in constructors for various NLP-relevant models, including several varieties recurrent neural nets like Tree LSTMs and encoder-decoders.
+DyNet includes built-in constructors for various NLP-relevant models, including several varieties of recurrent neural nets like Tree LSTMs and encoder-decoders.
 This makes it easy for the programmer to build many common kinds of models.
@@
 \ms{Many factors, no clear winner; }
+We implemented a simple language modeling task in three deep learning frameworks, and measured their performance when running on a GPU.
+We found that DyNet converged faster than TensorFlow or Theano, but analyzing performance on a fine-grained level showed that all frameworks used the underlying GPU resources with comparable efficiency.
+Thus, efficiency is only one of several factors researchers should consider in choosing a deep learning framework.
+Users should also consider ease of installation and use, level of support available, and the suitability of the framework for their particular task.
 

### 2016-12-14 paper/main.tex
paper proofread

--- 2016-12-14 paper/main.tex
@@
 \usepackage[utf8]{inputenc}
+
+% align enviroment with reduced spacing around it
+\usepackage{amsmath}
+\usepackage{etoolbox}
+\newcommand{\zerodisplayskips}{%
+  \setlength{\abovedisplayskip}{0.25em}
+  \setlength{\belowdisplayskip}{0.25em}
+  \setlength{\abovedisplayshortskip}{0.25em}
+  \setlength{\belowdisplayshortskip}{0.25em}}
+\appto{\normalsize}{\zerodisplayskips}
+
 \usepackage{emnlp16}
@@
 \begin{abstract}
-Neural networks are omnipresent in natural language processing. 
-We benchmark three popular Python frameworks (DyNet, TensorFlow and Theano) on a realistic NLP task: language modeling.
-\ms{Change if needed:} We find that DyNet is significantly faster on this task. However, performance wasn't the only bottleneck during development, and there are other variables to consider.
+Neural networks are omnipresent in natural language processing (NLP).
+We benchmark three popular Python frameworks (DyNet, TensorFlow, and Theano) on the standard NLP task of language modeling, and find that DyNet is significantly faster on this task.
+We also discuss other bottlenecks beyond performance, such as ease of use, that may impact the selection of a neural network framework.
 \end{abstract}
@@
 Several deep learning frameworks have been released in response to this, each claiming niche improvements over others, but it is not necessarily clear which frameworks are better suited for which types of applications.
-Previous comparisons may now be out-of-date (older versions; lack of newer frameworks) or were run against simpler tasks (which might have hidden other interesting performance characteristics).
+Previous comparisons betewen frameworks may now be out-of-date (older versions; lack of newer frameworks) or were run against simpler tasks (which might have hidden other interesting performance characteristics).
 
-In this project, we benchmark deep learning frameworks on a complex natural language processing task and rank them based on different performance metrics.
+In this project, we benchmark deep learning frameworks on a complex natural language processing (NLP) task and rank them based on different performance metrics.
 Specifically, we look at the following frameworks:
@@
 \end{itemize}
-In natural language processing (NLP), recurrent neural networks (RNNs) are of particular interest because they make use of the sequential nature of language.
+In NLP, recurrent neural networks (RNNs) are of particular interest because they make use of the sequential nature of language.
 Benchmarks on RNNs have been run before, but most tasks were benchmarked on toy tasks which do not necessarily reflect performance on real data.
@@
 generated random data to run through their RNN \cite{glample}, which is not realistic.
-In NLP, we usually turn words into vocabulary-sized one-hot vectors and then embed them (i.e.
-dimensionality reduction) as hidden-sized vectors.
-Sparsely learning the embedding matrix, which is usually in the order of $[20,000 \times 200]$, is likely to severely impact performance.
+A standard approach in NLP is to transform words into vocabulary-sized one-hot vectors and then embed them as hidden-sized vectors (i.e., dimensionality reduction).
+Sparse learning of the embedding matrix, which is usually in the order of $[20,000 \times 200]$, is likely to severely impact performance.
 
@@
 The task of language modeling is about estimating probabilities of particular sequences of words.
-It is used as an invaluable component in various real-world applications such as speech recognition, statistical machine translation, etc.
-More formally, if $s=\{START,x_1,x_2,...,x_n,STOP\}$ is an n-word sentence, language modelling will try to estimate, for each word $x_i$
-its probability given its history: $p(x_i|START,x_1,x_2,...,x_{i-1})$. 
-Historically, this task can be done rather simply using an l-th order Markov model, where $p(x_i|START,x_1,x_2,...,x_{i-1}) = p(x_i|x_{i-l},x_{i-l+1},...,x_{i-1})$.
-However, the Markov assumption does not need to be made with recurrent neural networks, as they have the capabilities to encode much longer histories.
-With RNNs, we normalize (using softmax) the output of a nonlinear function $f\left(p(x_i|START,x_1,x_2,...,x_{i-1})\right) =softmax\left(f(x_i,START,x_1,x_2,...,x_{i-1})\right)$.
+It is used as an invaluable component in various real-world applications such as speech recognition and statistical machine translation.
+More formally, if $s=\{START,x_1,x_2,...,x_n,STOP\}$ is an n-word sentence, language modelling will try to estimate
+the probability of each word $x_i$ given its history: $p(x_i|START,x_1,x_2,...,x_{i-1})$.
+
+Historically, this task was done by simply using an l-th order Markov model:
+\begin{align*}
+    p(x_i|START,x_1,x_2,...,x_{i-1}) = \\
+        p(x_i|x_{i-l},x_{i-l+1},...,x_{i-1})
+\end{align*}
+However, the Markov assumption does not need to be made with RNNs, as they are capable of encoding longer histories. With RNNs, we normalize (using softmax) the output of a nonlinear function $f$:
+\begin{align*}
+	f\left(p(x_i|START,x_1,x_2,...,x_{i-1})\right) = \\
+		softmax\left(f(x_i,START,x_1,x_2,...,x_{i-1})\right)
+\end{align*}
 
@@
 \subsection{Implementation \& Hyperparameters}
-We implemented the same recurrent neural language model in each framework.
-The model consists of a matrix of input word embeddings, which are passed through a RNN to get predicted word embeddings, which are compared via a loss function to a corresponding set of output embeddings.
+We implemented the same RNN language model in each framework.
+The model consists of a matrix of input word embeddings, which are passed through a RNN to get predicted word embeddings and then compared (with a loss function) to a corresponding set of output embeddings.
 All these elements are available in all frameworks (in Theano's case, the RNN is provided by the Keras wrapper).
 Our implementations are made available on GitHub.\footnote{\url{https://github.com/lucylin/neural-benchmarks}}
+
 In our experiments, we optimized for a standard language modeling objective: \textit{cross-entropy}.
@@
 Traditionally, neural models are trained until their performance on a \textit{development} set stops improving.
-Therefore, a full epoch of training (where all the training data is used) is usually followed by a predictive pass through development data.
+To determine if we should terminate, each full epoch of training (where all the training data is used) was followed by a predictive pass through the development data.
+
 \section{Benchmarks}
@@
 \subsection{Time}
-The first benchmark to look at is how long training takes. Table \ref{tab:timing} summarizes various breakdowns. \ms{Summarize results: ie DyNet is the best (so far) (2 sentences perhaps)}.
+The first benchmark to look at is how long training takes.
+Table \ref{tab:timing} summarizes various breakdowns\footnote{We were not able to complete a convergence run for the Theano implementation in time for this paper.
+Based on the epoch-specific measurements and comparison of losses, Theano's performance on this task is likely to be comparable to TensorFlow's performance; we are confident that it would not surpass DyNet on this task.}.
+Overall, DyNet seems to converge the fastest of the three; its epoch time is dramatically shorter, but it also took more epochs to converge.
+
 \begin{table}
@@
 Time					& DyNet 		& TensorFlow 	& Theano \\ \hline
-Convergence		& 26h10m 		& 35h25m 			& (TBD)\\
+Convergence		& 26h21m 		& 35h25m 			& * \\
 Train epoch 		& 2h20m		& 6h57m 			& 6h35m \\
@@
 \subsection{Kernel function usage}
+\label{subsec:kfunc}
 Using NVidia's built-in GPU profiler, \verb!nvprof!, we first looked into which kernel function calls were most heavily utilized by profiling training on a single batch.
-As can be seen in Table~\ref{tab:pcttime}, a large portion of the runtime is spent on matrix multiplication (specifically, a variant of \texttt{\detokenize{magma_lds_128_sgemm_kernel()}}) for all frameworks; this is expected given how neural networks work.
+As can be seen in Table~\ref{tab:pcttime}, a large portion of the runtime is spent on matrix multiplication (specifically, a variant of \texttt{\detokenize{magma_lds_128_sgemm_kernel()}}) for all frameworks.
+This is expected given that neural network operations are heavily dependent on matrix multiplies.
 
@@
 As discussed in section \ref{sec:benchmarks}, we found that DyNet was the fastest at performing our specific language modeling task.
-\lhl{TODO: flesh this out with a sentence or two more -- maybe something like ``DyNet seems to make certain optimizations that are advantageous in this setting''}
+DyNet seems to make certain optimizations that are advantageous in this setting, perhaps in the memory management choices it makes (as discussed in \ref{subsec:kfunc}).
+We also suspect (but could not empirically confirm based on metrics) that because DyNet dynamically creates computation graphs, it might be better able to adapt to different sentence lengths in a way that TensorFlow and Theano cannot.
 
-However, we note that the comparative performance between frameworks is likely dependent on the task, data set, choice of optimizer, and other parameters.
+Despite our performance findings, we note that the comparative performance between frameworks is likely dependent on the task, data set, choice of optimizer, and other parameters.
 Therefore, these results may not extrapolate to other settings, and if tuning training or prediction performance is a serious consideration, we recommend benchmarking on the specific task at hand.
@@
 \paragraph{TensorFlow}
+TensorFlow provides very extensive installation guidelines, and pip install takes care of installing it and its dependencies.
+However, one inconvenience is that it requires separate installations for CPU and GPU usage, which can cause issues if not using a virtual environment of some sort.
 
@@
 \subsubsection{Ease of use}
-\ms{TF pro: Google has money and resources.}
-\lhl{TODO: all the things that TensorFlow/DyNet support out of the box.
-comment somewhere about documentation, debuggability}
 
 \paragraph{DyNet}
-DyNet includes built-in constructors for various NLP-relevant models, including several varieties recurrent neural nets like Tree LSTMs and encoder-decoders.
-THis makes it easy for the programmer to build many common kinds of models.
-DyNet also allows for dynamic computation graph, allowing nodes (representing parameter tensors) to be combined ``on the fly'' to adapt to, for example, data sequences of different lengths.
+DyNet includes built-in constructors for various NLP-relevant models, including several varieties of RNNs (e.g., tree LSTMs, encoder-decoders).
+This makes it easy for the programmer to build many common kinds of models.
+DyNet also allows for dynamic computation graphs, allowing nodes (representing parameter tensors) to be combined ``on the fly'' to adapt to, for example, data sequences of different lengths.
 While our neural language model used a static computation graph, DyNet made graph construction easy.
@@
 
-\paragraph{TensorFlow} \textless TensorFlow stuff goes here \textgreater
+\paragraph{TensorFlow} TensorFlow is relatively easy to use, as it comes with a myriad of tutorials and examples.
+TensorFlow's large userbase, polyvalence, and large amount of support from its creator (Google) have made it a very solid option for developers.
+
+There is plenty of support for RNNs and many other standard models.
+However, writing the computation graph requires some amount of know-how, since there is a lot of TensorFlow-specific syntax, and debugging the symbolic graph can be difficult since there are no values.
+TensorFlow also sometimes requires a decent amount of boilerplate code; for example, the computation graph being statically constructed requires that sequences or mini-batches be padded, which can make RNN computation somewhat cumbersome.
 
 \paragraph{Theano}
-In contrast, Theano is ``lower-level'' in that it requires that the programmer define the shared variables, propagation steps, cost functions, and so on manually.
+In contrast to DyNet and TensorFlow, Theano is ``lower-level'' in that it requires that the programmer define the shared variables propagation steps, and so on manually.
 Theano therefore offers a great deal of control and flexibility in model implementation, which is a benefit if one is implementing neural net layers not supported by other frameworks.
@@
 However, this also presents a steeper learning curve (despite the extensive documentation) and a much higher prototyping/development cost.
-We originally attempted to implement the Theano language model just using Theano and found the learning/debugging costs to be high.
-Because our language model consists of very standard neural net components, we instead switched to using Keras, which can use Theano under the covers but supplies an easier API with built-in RNN components.
+We originally attempted to implement the Theano language model using just Theano and found the learning/debugging costs to be high.
+Because our language model consists of very standard neural net components, we instead switched to using Keras, which supplies an API with built-in RNN components implemented using Theano.
 
 \section{Conclusion}
-\ms{Many factors, no clear winner; }
+We implemented a simple language modeling task in three deep learning frameworks and measured their performance when run on a GPU.
+We found that DyNet converged faster than TensorFlow or Theano, but performance analysis on a fine-grained level showed that all frameworks used the underlying GPU resources with comparable efficiency.
+Through our experience implementing language models in these frameworks, we also found that efficiency is only one of several factors researchers should consider in choosing a deep learning framework.
+Users should also consider ease of installation and use, level of support available, and the suitability of the framework for their particular task.
 
 \section*{Acknowledgments}
-We want to thank Emily Furst for helping us with GPU profiling, specifically understanding the CUDA Profiler documentation.
+We want to thank Emily Furst for her help with GPU profiling (specifically, understanding the CUDA Profiler documentation).
 \newpage

### 2016-12-14 theano/defaults.py
checking in keras/theano code

@@ -1,0 +1,11 @@
+# Default values of things.
+
+LEARNING_RATE = 0.003
+HIDDEN_SIZE = 256
+BATCH_SIZE = 25
+MAX_EPOCH = 50
+
+# TensorFlow/Dynet defaults that we need to define manually here
+ADAMOPT_BETA1 = 0.9
+ADAMOPT_BETA2 = 0.999
+ADAMOPT_EPSILON = 1e-8

### 2016-12-14 theano/lm.py

@@ -1,0 +1,206 @@
+import itertools
+import logging
+import numpy as np
+import time
+import defaults
+
+from keras.utils.np_utils import to_categorical
+from nn import RNNLM
+
+UNK = 0
+BOS = 1
+EOS = 2
+
+VOCAB_SIZE = 35642+1 # words + OOV + BOS + EOS + 1 to correct for UNK!=0
+
+class LanguageModel:
+    '''
+    Basic language model using Keras-based LSTM implementation.
+    '''
+
+    def __init__(self):
+        self._model = None
+        self._log = logging.getLogger(self.__class__.__name__)
+
+    def configure_logger(self, level=logging.DEBUG, write_file=False):
+        fmt = '%(asctime)s %(name)s.%(funcName)-20s %(levelname)-8s \
+               %(message)s'
+
+        kwargs = {
+            'level': level,
+            'format': fmt,
+            'datefmt': '%m-%d %H:%M',
+        }
+
+        if write_file:
+            log_path = 'theano.log'  # TODO: better.
+
+            kwargs['filename'] = log_path
+            kwargs['filemode'] = 'w'
+
+        logging.basicConfig(**kwargs)
+        self._log = logging.getLogger(self.__class__.__name__)
+
+    def _load_sequences(self, data_path):
+        self._log.info('Loading data from {0}...'.format(data_path))
+
+        with open(data_path, 'r') as f:
+            seqs = [[int(x) for x in line.strip().split()] for line in f]
+
+        # +1 since 0-indexed
+        vocab_size = max(max(s) for s in seqs) + 1
+
+        self._log.debug('Data loaded!')
+        return seqs, vocab_size
+
+    '''Training-specific functions.'''
+
+    def train(self, train_path, output_path=None,
+              learning_rate=defaults.LEARNING_RATE,
+              hidden_size=defaults.HIDDEN_SIZE, batch_size=defaults.BATCH_SIZE,
+              max_epoch=defaults.MAX_EPOCH):
+
+        self._log.info('Enter: train({0})'.format(locals()))
+
+        new_unk = VOCAB_SIZE + 1
+
+        def get_data(data_path):
+            seqs, vocab_size = self._load_sequences(data_path)
+            max_seq_len = max(len(s) for s in seqs)
+
+            if vocab_size > VOCAB_SIZE:
+                raise Exception('unexpectedly large vocab size')
+
+            return self._format_data(seqs, new_unk), max_seq_len
+
[…]

### 2016-12-14 theano/nn.py

@@ -1,0 +1,52 @@
+import numpy as np
+import theano
+import theano.tensor as T
+
+import defaults
+
+from keras.callbacks import EarlyStopping
+from keras.models import Sequential
+from keras.layers.core import Dense, Activation
+from keras.layers.embeddings import Embedding
+from keras.layers.recurrent import LSTM
+from keras.layers.wrappers import TimeDistributed
+from keras.optimizers import Adam
+
+class RNNLM:
+    def __init__(self, hidden_size, vocab_size, batch_size, learning_rate,
+                 max_seq_len):
+
+        self.hidden_size = hidden_size
+        self.vocab_size = vocab_size
+        self.batch_size = batch_size
+        self.learning_rate = learning_rate
+        self.max_seq_len = max_seq_len
+
+        self._build_graph()
+
+    def _build_graph(self):
+        model = Sequential()
+
+        input_shape = (self.batch_size, self.max_seq_len)
+        model.add(Embedding(self.vocab_size, self.hidden_size, mask_zero=True,
+                  batch_input_shape=input_shape))
+
+        model.add(LSTM(self.hidden_size, return_sequences=True, stateful=True,
+                  consume_less='gpu'))   # change as needed
+
+        # unemb/softmax layer
+        model.add(TimeDistributed(Dense(self.vocab_size)))
+        model.add(Activation('softmax'))
+
+        opt = Adam(lr=self.learning_rate)
+        model.compile(loss='categorical_crossentropy', optimizer=opt,
+                      metrics=['categorical_crossentropy'])
+
+        self.model = model
+
+    def train_batch(self, data, labels):
+        return self.model.fit(data, labels, nb_epoch=1,
+                              batch_size=self.batch_size, shuffle=False)
+
+    def validate(self, data, labels):
+        return self.model.evaluate(data, labels, batch_size=self.batch_size)

### 2016-12-14 theano/run.py

@@ -1,0 +1,65 @@
+import argparse
+import logging
+
+# Module-specific
+import defaults
+from lm import LanguageModel
+
+
+def get_argparser():
+    """Return an ArgumentParser with default arguments."""
+
+    p = argparse.ArgumentParser()
+
+    p.add_argument('--DEBUG', action='store_true', help="Log on DEBUG mode")
+
+    p.add_argument("--data_path", metavar="PATH",
+                   help="path where the data is stored as a Pandas DataFrame")
+    p.add_argument("--test", metavar="PATH",
+                   help="tests a model loaded from a specified path (uses \
+                         gold standard inputs)")
+    p.add_argument("--train", metavar="PATH",
+                   help="trains a model and saves it to a specified path")
+
+    # Hyper parameters ish
+    p.add_argument("--learning_rate", metavar="LR", type=float,
+                   default=defaults.LEARNING_RATE,
+                   help="Learning rate for AdamOptimizer.")
+    p.add_argument("--hidden_size", dest="hidden_size", type=int,
+                   default=defaults.HIDDEN_SIZE,
+                   help="Size of tensors in hidden layer.")
+    p.add_argument("-b", "--batch_size", dest="batch_size", type=int,
+                   default=defaults.BATCH_SIZE,
+                   help="Batch size to pass through the graph.")
+    p.add_argument("--max_epoch", dest="max_epoch", type=int,
+                   default=defaults.MAX_EPOCH,
+                   help="Determines how many passes throught the training \
+                         data to do")
+
+    return p
+
+
+def main():
+    p = get_argparser()
+    args = p.parse_args()
+
+    lm = LanguageModel()
+    lm.configure_logger(level=logging.DEBUG if args.DEBUG else logging.INFO,
+                        write_file=True)
+
+    if args.train and args.data_path:
+        lm.train(args.data_path, output_path=args.train,
+                 learning_rate=args.learning_rate,
+                 hidden_size=args.hidden_size,
+                 batch_size=args.batch_size, max_epoch=args.max_epoch)
+
+    elif args.test and args.data_path:
+        lm.predict(args.test, args.data_path)
+
+    else:
+        # Well, this is silly.
+        p.print_help()
+        exit(2)
+
+if __name__ == '__main__':
+    main()

### 2016-12-14 paper/main.tex
slightly better

@@
 \paragraph{Theano}
-In contrast to DyNet and TensorFlow, Theano is ``lower-level'' in that it requires that the programmer define the shared variables propagation steps, and so on manually.
-Theano therefore offers a great deal of control and flexibility in model implementation, which is a benefit if one is implementing neural net layers not supported by other frameworks.
+In contrast to DyNet and TensorFlow, Theano is ``lower-level'' in that it requires that the programmer define the shared variables propagation steps, and so on manually.\footnote{An example of the relative implementation complexity can be seen in code for the LSTM Theano tutorial at \url{http://deeplearning.net/tutorial/lstm.html}.} Theano therefore offers a great deal of control and flexibility in model implementation, which is a benefit if one is implementing neural net layers not supported by other frameworks.
 
@@
 We found that DyNet converged faster than TensorFlow or Theano, but performance analysis on a fine-grained level showed that all frameworks used the underlying GPU resources with comparable efficiency.
+
 Through our experience implementing language models in these frameworks, we also found that efficiency is only one of several factors researchers should consider in choosing a deep learning framework.

### 2016-12-14 paper/main.tex
typo fixes

@@
 Several deep learning frameworks have been released in response to this, each claiming niche improvements over others, but it is not necessarily clear which frameworks are better suited for which types of applications.
-Previous comparisons betewen frameworks may now be out-of-date (older versions; lack of newer frameworks) or were run against simpler tasks (which might have hidden other interesting performance characteristics).
+Previous comparisons between frameworks may now be out-of-date (older versions; lack of newer frameworks) or were run against simpler tasks (which might have hidden other interesting performance characteristics).
 
@@
 The frameworks handle computation graphs differently.
-Theano and Tensorflow perform ``static'' declaration of computation graphs: first, the graph architecture is specified, and then data is run through the compiled graph to train or make predictions.
+Theano and TensorFlow perform ``static'' declaration of computation graphs: first, the graph architecture is specified, and then data is run through the compiled graph to train or make predictions.
 In contrast, DyNet does ``dynamic'' declaration of computation graphs; that is, the computation graph is defined on-the-fly as operations are executed.
